### 2022-11-23 ms.rmd
Adding template of MS in Rmarkdown

---
title: ""
author: et al. 
date: "`r Sys.Date()`"
bibliography: ./bib/refs.bib
csl: ./bib/ecology-letters.csl
output: 
  bookdown::word_document2:
    toc: no
    toc_depth: 6
    number_sections: false
    reference_docx: ./bib/template.docx
  bookdown::html_document2:
    code_folding: hide
    number_sections: no
    toc: yes
    toc_depth: 6
    toc_float: yes
    bibliography: refs.bib
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, tidy = TRUE, fig.width = 10)
## numbers >= 10^5 will be denoted in scientific notation,		  ## numbers >= 10^5 will be denoted in scientific notation,
  ## and rounded to 2 digits		  ## and rounded to 2 digits
  options(digits = 2)

##################################
# Clean workspace
##################################
    rm(list=ls())

##################################
# Loading packages & Functions
##################################
    pacman::p_load(tidyverse,  metafor, flextable, raster, rasterVis,  patchwork, RColorBrewer, ggtree, ape, phytools, viridis, MCMCglmm, brms, latex2exp, orchaRd, multcomp)

```

## References

### 2022-11-24 ms.rmd
Adding tree data

@@
 ##################################
-    pacman::p_load(tidyverse,  metafor, flextable, raster, rasterVis,  patchwork, RColorBrewer, ggtree, ape, phytools, viridis, MCMCglmm, brms, latex2exp, orchaRd, multcomp)
+    pacman::p_load(tidyverse,  metafor, flextable,  patchwork, ggtree, ape, phytools, viridis, MCMCglmm, brms, latex2exp, orchaRd, multcomp)
+
+```
+
+```{r, tree}
+# Bring in the fly tree
+tree <- read.tree("./phylogeny/drosi.tre")
+
+# Plot the tree
+plot(tree, cex = 0.2)
+
+# Ultrametric?
+is.ultrametric(tree) # nope
+
+# Grab phylo matrix
+phy_vcv <- vcv(tree)
 

### 2022-11-24 R/Analyses_CTmax_z.R
Adding in some caluclations with Enrico

@@ -63,7 +63,7 @@ text(2,-12,substitute("LT"[50]))
 mtext("Temperature (ºC)",2,line=1,outer=TRUE,cex=0.9)
 mtext("Time (log-scale)",1,line=1.2,outer=TRUE,adj=0.07,cex=0.9)
 
-# Plot CTmin versus z' 
+# Plot CTmin versus z'
 par(mar=c(2,5,4,1))
 coef <- coef(lm(data$Ctmin ~ data$z + as.factor(data$LT)))
 data$Ctmin.trans <- ifelse(data$LT==50,data$Ctmin,data$Ctmin - coef[3])
@@ -117,7 +117,7 @@ arrows(2.7,23.7,3.3,23.7,xpd=TRUE,length=0.1)
 
 # Total number of measurements estimated by Enrico on June 2019 for scaling study with Nacho:
 n <- c(4,4,5,5,3,7,7,4,4,4,4,4,4,7,7,7,4,6,5,4,4,4,5,6,5,8,4,3,7,7,7,3,5,6,6,9,3,6,3,3,4,9,4,3,5,5,5,6,5,5,4,3,5,7,7,7,4,3,3,6,6,3)
- 
+
 
 # ArmstrongEA2009_Table_5
 t1 <- c(120,30,3,1.5)
@@ -125,9 +125,23 @@ t2 <- c(80,25,6,1)
 t3 <- c(50,12,3,0.5)
 t4 <- c(70,20,4,1)
 T <- c(44,46,48,50)
+#################################################
+# EXAMPLE CALCULATION WITH ENRICO
+#################################################
 
-# ArmstrongEA2009_Table_6
t1 <- c(60,20,5,1.5)
t2 <- c(130,25,4,1.5)
-t3 <- c(70,15,2.5,1.0)
t4 <- c(55,15,5,1.5)
+mod_arstrong1 <- lm(log10(t1)~ T) # Need to have log10 time on y because it's the only thing that varies, so residual
+summary(mod_arstrong1)
+    z1 = 1/coef(mod_arstrong1)[2]
+ se_z1 = sqrt(diag(vcov(mod_arstrong1))[2])^0.5 # convert se from original scale to new scale
+
+ctmax1 = coef(mod_arstrong1)[1] / coef(mod_arstrong1)[2]
[…]

### 2022-11-24 ms.rmd
Adding tentative model

@@
 
+```{r}
+## Models
+   ctmax <- mv(ctmax | se(sqrt(ctmax_se)) ~ 1 + (1|q|gr(phylo = phy_vcv)) + (1|spp) + (1|obs)) + gaussian()
+    cfts <- mv( cfts | se(sqrt(cfts_se))  ~ 1 + (1|q|gr(phylo = phy_vcv)) + (1|spp)) + gaussian()
+   tdt_i <- mv(tdt_i |  mi() ~ 1 + (1|q|gr(phylo = phy_vcv)) + (1|spp)) + gaussian()
+   tdt_z <- mv(tdt_z |  mi() ~ 1 + (1|q|gr(phylo = phy_vcv)) + (1|spp)) + gaussian()
+
+       mv_brms_model <- brms::brm(ctmax + tdt_i + tdt_z + cfts, 
+                        warmup = 1000, iter = 4000, thin = 5, 
+                        control = list(adapt_delta = 0.95, max_treedepth = 15), 
+                        data = data, data2 = phy_vcv)
+
+```
+
 ## References

### 2022-11-24 R/Analyses_CTmax_z.R
Another example of calculation

@@ -168,6 +168,8 @@ T <- c(41,42,43,44,45,46)
 t <- c(31,15,8)
 T <- c(39,40,41)
 
+mod_feder <- lm(log10(t) ~ T)
+
 #Brett1956 Fig 2
 t1 <- c(23.2,72.6,143.2,271.6,853.1),
 T1 <- c(34.5,34,33.5,33,32.5),
@@ -187,6 +189,10 @@ T7 <- c(27,26.5,26,25),,
 #CooperEA2008_Fig2
 t <- c(27.5,19.2,8.5,10.2,10.2,5.7,1.4,1.8,1.6)
 T <- c(38,38.5,39,39.5,40,40.5,41,41.5,42)
+
+mod_cooper <- lm(log10(t) ~ T)
+z_cooper <- 1 / coef(mod_cooper)[2]
+ctmax_cooper <- coef(mod_cooper)[1] / coef(mod_cooper)[2]
 
 #QuinnEA1994_Table1
 t <- c(24,48,96)*60

### 2022-11-24 R/func.R
added tree check function

@@ -1,0 +1,21 @@
+#Tree checks function
+tree_checks <- function(data, tree, dataCol, type = c("checks", "prune")){
+  type = match.arg(type)
+  # How many unique species exist in data and tree
+  Numbers <- matrix(nrow = 2, ncol = 1)
+  Numbers[1,1] <- length(unique(data[,dataCol]))
+  Numbers[2,1] <- length(tree$tip.label)
+  rownames(Numbers)<- c("Species in data:", "Species in tree:")
+  # Missing species or species not spelt correct
+  species_list1= setdiff(sort(tree$tip.label), sort(unique(data[,dataCol])))
+  species_list2= setdiff(sort(unique(data[,dataCol])), sort(tree$tip.label) )
+  if(type == "checks"){
+    return(list(SpeciesNumbers = data.frame(Numbers),
+                Species_InTree_But_NotData=species_list1,
+                Species_InData_But_NotTree=species_list2))
+  }
+  if(type == "prune"){
+    if(length(species_list2) >=1) stop("Sorry, you can only prune a tree when you have no taxa existing in the data that are not in the tree")
+    return(ape::drop.tip(tree, species_list1))
+  }
+}

### 2022-11-24 R/Colinet.R
fixed a few spelling errors

@@ -1,0 +1,12 @@
+
+T <- c(37,35,34,33,32)
+t1 <- c(3.232,7.790,8.321,11.132,24.281)*60
+t2 <- c(3.866,15.358,22.070,29.810,32.785)*60
+
+a <- lm(log10(t1) ~ T)
+coef(a)[1]/coef(a)[2]
+1/coef(a)[2]
+
+b <- lm(log10(t2) ~ T)
+coef(b)[1]/coef(b)[2]
+1/coef(b)[2]

### 2022-11-24 ms.rmd
New tree

--- 2022-11-24 ms.rmd
@@
 # Bring in the fly tree
-tree <- read.tree("./phylogeny/drosi.tre")
+tree <- read.tree("./phylogeny/drosi2.tre")
 
@@
 is.ultrametric(tree) # nope
+tree2 <- force.ultrametric(tree, method = "extend")
 
+# Grab out data
+   data <- read.csv("./data/drosi_data.csv")
+
+# functions
+source("./R/func.R")
+t <- tree_checks(tree2, data = data, dataCol = "species", type = "check")
+data <- data %>% mutate(species = gsub(" ", "", species),
+                        phylo = species,
+                        obs = 1:n()) %>% filter(!species %in% t$Species_InData_But_NotTree) 
+tree3 <- tree_checks(tree2, data = data, dataCol = "species", type = "prune")
+tree_checks(tree3, data = data, dataCol = "species", type = "check")
 # Grab phylo matrix
-phy_vcv <- vcv(tree)
+phy_vcv <- vcv(tree3, cor = FALSE)
 
@@
 
+```{r}
+
+## Model(s)
+   ctmax <- bf(ctmax | se(sqrt(ctmax_se)) ~ 1 + (1|q|gr(phylo, cov = phy_vcv)) + (1|species) + (1|obs)) + gaussian()
+    cfts <- bf( cfts |  mi() ~ 1 + (1|q|gr(phylo, cov = phy_vcv)) + (1|species)) + gaussian()
+   #tdt_i <- bf(tdt_i |  mi() ~ 1 + (1|q|gr(phylo = phy_vcv)) + (1|spp)) + gaussian()
+   tdt_z <- bf(tdt_z |  mi() ~ 1 + (1|q|gr(phylo, cov = phy_vcv)) + (1|species)) + gaussian()
+
+       mv_brms_model <- brms::brm(ctmax + tdt_z + cfts, 
+                        warmup = 1000, iter = 4000, thin = 5, 
+                        control = list(adapt_delta = 0.95, max_treedepth = 15), cores = 4,
+                        data = data, data2 = list(phy_vcv = phy_vcv))
+
+#### Try MCMCglmm which I expect will be MUCH faster. Deals with missing data using data augmentation
+
+prior = list(R = list(V = diag(3), nu=1.002), 
+             G = list(G1=list(V = diag(3), nu = 1.002)))
+
+data$animal = data$phylo
+
+mcmcglmm_model <- MCMCglmm::MCMCglmm(cbind(ctmax, cfts, tdt_z) ~ -1 + trait, random = ~us(trait):animal, rcov = ~us(trait):units ,
+                                     pedigree=tree3,
+                                     burnin=1000, nitt = 300000, thin = 5, family = rep("gaussian", 3), prior = prior, 
+                                    pl = TRUE, pr = TRUE, data = data)
+saveRDS(mcmcglmm_model, file = "./output/model/mcmcglmm_model")
+```
+
 ## References

### 2022-11-26 ms.rmd
Added node labels = NULL and added a models with scaled variables to make it easier to check correlation at phlo level

--- 2022-11-24 ms.rmd
@@
 is.ultrametric(tree) # nope
+tree2 <- force.ultrametric(tree, method = "extend")
 
+# Grab out data
+   data <- read.csv("./data/drosi_data.csv")
+
+# functions
+source("./R/func.R")
+t <- tree_checks(tree2, data = data, dataCol = "species", type = "check")
+data <- data %>% mutate(species = gsub(" ", "", species),
+                        phylo = species,
+                        obs = 1:n()) %>% filter(!species %in% t$Species_InData_But_NotTree) 
+tree3 <- tree_checks(tree2, data = data, dataCol = "species", type = "prune")
+tree_checks(tree3, data = data, dataCol = "species", type = "check")
 # Grab phylo matrix
-phy_vcv <- vcv(tree)
+phy_vcv <- vcv(tree3, cor = FALSE)
+tree3$node.label <- NULL
+
+```
+
+```{r}
+
+#### Try MCMCglmm which I expect will be MUCH faster. Deals with missing data using data augmentation
+rerun = FALSE
+if(rerun){
+  prior = list(R = list(V = diag(3), nu=1.002), 
+             G = list(G1=list(V = diag(3), nu = 1.002)))
+
+data$animal = data$phylo
+
+mcmcglmm_model <- MCMCglmm::MCMCglmm(cbind(ctmax, cfts, tdt_z) ~ -1 + trait, random = ~us(trait):animal, rcov = ~us(trait):units ,pedigree=tree3,burnin=1000, nitt = 300000, thin = 5, family = rep("gaussian", 3), prior = prior, pl = TRUE, pr = TRUE, data = data)
+
+# We'll also z-transofmr to make the coprrelation stuff easier to calculate. 
+data <- data %>% mutate(z_ctmax = scale(ctmax),
+                         z_cfts = scale(cfts),
+                        z_tdt_z = scale(tdt_z))
+
+mcmcglmm_model_z_trans <- MCMCglmm::MCMCglmm(cbind(z_ctmax, z_cfts, z_tdt_z) ~ -1 + trait, random = ~us(trait):animal, rcov = ~us(trait):units, pedigree=tree3,burnin=1000, nitt = 300000, thin = 5, family = rep("gaussian", 3), prior = prior, pl = TRUE, pr = TRUE, data = data)
+
+saveRDS(mcmcglmm_model, file = "./output/model/mcmcglmm_model")
+saveRDS(mcmcglmm_model_z_trans, file = "./output/model/mcmcglmm_model_z_trans")
+} else{
+  mcmcglmm_model <- readRDS("./output/model/mcmcglmm_model")
+}
+
+# Extract estimates
+sol <- mcmcglmm_model$Sol
+vcv <- mcmcglmm_model$VCV
+
+# Extract the estmated / imputaed Z for each species, but drop internal ancestral nodes
+z <- sol[, grep("traittdt_z.animal", colnames(sol))]
+z <- z[,-grep("Node", colnames(z))]
+head(z)
+
+# Clean up names so they are just species names
 

### 2022-11-26 ms.rmd
OK, we've got a no phylo model too so that we can be sure phylogeny makes a difference

@@
 
+# Ditch phylogeny to see if there is support for including it or not. DIC is WAY lower for phylogenetic model, so good to include it.
+prior2 = list(R = list(V = diag(3), nu=1.002))
+mcmcglmm_model_z_trans_nophylo <- MCMCglmm::MCMCglmm(cbind(z_ctmax, z_cfts, z_tdt_z) ~ -1 + trait, rcov = ~us(trait):units, pedigree=tree3,burnin=1000, nitt = 300000, thin = 5, family = rep("gaussian", 3), prior = prior2, pl = TRUE, pr = TRUE, data = data)
+
 saveRDS(mcmcglmm_model, file = "./output/model/mcmcglmm_model")
 saveRDS(mcmcglmm_model_z_trans, file = "./output/model/mcmcglmm_model_z_trans")
+saveRDS(mcmcglmm_model_z_trans_nophylo, file = "./output/model/mcmcglmm_model_z_trans_nophylo")
 } else{
   mcmcglmm_model <- readRDS("./output/model/mcmcglmm_model")
+  mcmcglmm_model_z_trans <- readRDS("./output/model/mcmcglmm_model_z_trans")
+mcmcglmm_model_z_trans_nophylo <- readRDS("./output/model/mcmcglmm_model_z_trans_nophylo")
 }

### 2022-11-28 ms.rmd
Doing some re-organisation and checking imputed values to real ones for cross valication

@@
   tree_checks(tree3, data = data, dataCol = "species", type = "check")
+
   # Grab phylo matrix
@@
 
-```{r}
+```{r, models}
 
-#### Try MCMCglmm which I expect will be MUCH faster. Deals with missing data using data augmentation
+### MCMCglmm Bayesian models. Deals with missing data using data augmentation
+
 rerun = FALSE
@@
 
-# We'll also z-transofmr to make the coprrelation stuff easier to calculate. 
+  # We'll also z-transform to make the correlation easier to calculate. 
   data <- data %>% mutate(z_ctmax = scale(ctmax),
@@
 
-# Extract estimates
+```
+
+```{r, posteriors}
+
+# Extract posterior estimates
   sol <- mcmcglmm_model$Sol
@@
 
-# Extract the estmated / imputaed Z for each species, but drop internal ancestral nodes
+# Extract the estimated / imputed Z for each species, but drop internal ancestral nodes
   z <- sol[, grep("traittdt_z.animal", colnames(sol))]
@@
 # Clean up names so they are just species names
+  colnames(z) <- gsub("traittdt_z.animal.", "", colnames(z))
+
+# Have a look at how well the model gets the actual data correct.
+  z_known <- data %>% filter(!is.na(tdt_z)) %>% dplyr::select(species, tdt_z) %>% distinct()
+    z_est <- apply(z, 2, function(x) median(x))[which(names(z_est) %in% z_known$species)]
 

### 2022-11-28 ms.rmd
Added phylogenetic signal calculation. Need to figure out the incoperation of covariance at animal and units level into this calculation as they seem awfully high phylogenetic H2's

@@
 
+# Check phylogeneic signal. Need to think about the covariance here. 
+    # CTMax - Yes
+      phylo_ctmax <- vcv[,"traitctmax:traitctmax.animal"]
+      units_ctmax <- vcv[,"traitctmax:traitctmax.units"]
+
+      h2_ctmax = phylo_ctmax / (phylo_ctmax + units_ctmax)
+      posterior.mode(h2_ctmax)
+      HPDinterval(h2_ctmax)
+
+    # CFL - Yes
+      phylo_cfls <- vcv[,"traitcfts:traitcfts.animal"]
+      units_cfls <- vcv[,"traitcfts:traitcfts.units"]
+
+      h2_cfls = phylo_cfls / (phylo_cfls + units_cfls)
+      posterior.mode(h2_cfls)
+      HPDinterval(h2_cfls)
+
+    # z - Yes
+      phylo_z <- vcv[,"traittdt_z:traittdt_z.animal"]
+      units_z <- vcv[,"traittdt_z:traittdt_z.units"]
+
+      h2_z = phylo_z / (phylo_z + units_z)
+      posterior.mode(h2_z)
+      HPDinterval(h2_z)
 ```

### 2022-11-28 ms.rmd
Ok, you need to remember that these are BLUPS, deviations form intercept (duh) so one needs to add in the mean intercept.

@@
 # Extract the estimated / imputed Z for each species, but drop internal ancestral nodes
-  z <- sol[, grep("traittdt_z.animal", colnames(sol))]
-  z <- z[,-grep("Node", colnames(z))]
-  head(z)
+  z_blups <- sol[, grep("traittdt_z.animal", colnames(sol))]
+  z_blups <- z[,-grep("Node", colnames(z))]
+  head(z_blups)
+
+# Remember that these are BLUPs. They are deviations from the intercept. So, to get back to the z we need to add the intercept to each of them.
+  z <- z_blups + mean(sol[, "traittdt_z"])
 

### 2022-11-28 ms.rmd
Fixed a few things and checks the posterios are matching the real data..which they are. We nee dto dos ome cross validation now - maybe drop one species, impute it and compare it to it's real data. Do this for each species.

@@
   z_blups <- sol[, grep("traittdt_z.animal", colnames(sol))]
-  z_blups <- z[,-grep("Node", colnames(z))]
+  z_blups <- z_blups[,-grep("Node", colnames(z_blups))]
   head(z_blups)
@@
 # Remember that these are BLUPs. They are deviations from the intercept. So, to get back to the z we need to add the intercept to each of them.
-  z <- z_blups + mean(sol[, "traittdt_z"])
+  z <- z_blups + sol[, "traittdt_z"]
 
@@
   z_known <- data %>% filter(!is.na(tdt_z)) %>% dplyr::select(species, tdt_z) %>% distinct()
-    z_est <- apply(z, 2, function(x) median(x))[which(names(z_est) %in% z_known$species)]
+    z_est <- apply(z, 2, function(x) median(x))
+    z_est <- z_est[which(colnames(z) %in% z_known$species)]
+    z_est <- data.frame(z_est)
+    z_est$species <- row.names(z_est)
+  check_z <- left_join(z_known, z_est, by = "species")
 

### 2023-09-19 ms.qmd
switch to qmd

--- 2022-11-28 ms.rmd
@@
 ##################################
-    pacman::p_load(tidyverse,  metafor, flextable,  patchwork, ggtree, ape, phytools, viridis, MCMCglmm, brms, latex2exp, orchaRd, multcomp)
+    pacman::p_load(tidyverse,  metafor, flextable,  patchwork, ggtree, ape, phytools, viridis, MCMCglmm, brms, latex2exp, orchaRd, multcomp, bayesplot)
 
@@
 
+```{r, posteriorsz}
+# We'll now plot the posterior distributions for each species.Obviously our 11 species with known data will have a much tighter distribution.
+You can see that in the plot.
+But the imputations are our 'best" guess based on the strong phylogenetic signal in the data.
+We can also always propagate the uncertainty across any subsequent analysis (like for example our micro-climate models)
+  mcmc_areas(z[,1:15],prob = 0.8)
+```
+
 ## References

### 2023-09-19 ms/ms.qmd
MOved docs to folder to clean up

--- 2023-09-19 ms.qmd
@@
 ---
-title: ""
-author: et al. 
-date: "`r Sys.Date()`"
-bibliography: ./bib/refs.bib
-csl: ./bib/ecology-letters.csl
-output: 
-  bookdown::word_document2:
-    toc: no
-    toc_depth: 6
-    number_sections: false
-    reference_docx: ./bib/template.docx
-  bookdown::html_document2:
-    code_folding: hide
-    number_sections: no
-    toc: yes
-    toc_depth: 6
-    toc_float: yes
-    bibliography: refs.bib
+title: "Box 1 - TDT Curves"
+bibliography: ../bib/refs.bib
+csl: ../bib/nature.csl
+format:
+  docx:
+    reference-doc: ../bib/template.docx
 editor_options: 
   chunk_output_type: console
+execute:
+  freeze: auto  # re-render only when source changes
+  cache: false
+  echo: false
+  warning: false
+  error: false
+  include: true
+crossref:
+  fig-title: 'Figure'
+  fig-labels: arabic
+  title-delim: "-"
+  fig-prefix: "Figure"
+  tbl-prefix: "Table"
 ---
@@
 ##################################
-    pacman::p_load(tidyverse,  metafor, flextable,  patchwork, ggtree, ape, phytools, viridis, MCMCglmm, brms, latex2exp, orchaRd, multcomp, bayesplot)
+    pacman::p_load(tidyverse,  metafor, flextable,  patchwork, ggtree, ape, phytools, viridis, MCMCglmm, brms, latex2exp, orchaRd, multcomp, bayesplot, here)
 
@@
 # Bring in the fly tree
-  tree <- read.tree("./phylogeny/drosi2.tre")
+  tree <- read.tree(here::here("phylogeny/", "drosi2.tre"))
 
@@
 # Grab out data
-   data <- read.csv("./data/drosi_data.csv")
+   data <- read.csv(here::here("data", "drosi_data.csv"))
 
 # functions
-  source("./R/func.R")
+  source(here::here("R", "func.R"))
   t <- tree_checks(tree2, data = data, dataCol = "species", type = "check")
@@
 
-  saveRDS(mcmcglmm_model, file = "./output/model/mcmcglmm_model")
-  saveRDS(mcmcglmm_model_z_trans, file = "./output/model/mcmcglmm_model_z_trans")
-  saveRDS(mcmcglmm_model_z_trans_nophylo, file = "./output/model/mcmcglmm_model_z_trans_nophylo")
+  saveRDS(mcmcglmm_model, file =  here::here("output/model", "mcmcglmm_model"))
+  saveRDS(mcmcglmm_model_z_trans, file = here::here("output/model", "mcmcglmm_model_z_trans")) 
+  saveRDS(mcmcglmm_model_z_trans_nophylo, file =  here::here("output/model", "mcmcglmm_model_z_trans_nophylo")) 
 } else{
-                mcmcglmm_model <- readRDS("./output/model/mcmcglmm_model")
-        mcmcglmm_model_z_trans <- readRDS("./output/model/mcmcglmm_model_z_trans")
-mcmcglmm_model_z_trans_nophylo <- readRDS("./output/model/mcmcglmm_model_z_trans_nophylo")
+                mcmcglmm_model <- readRDS(here::here("output/model/", "mcmcglmm_model"))
+        mcmcglmm_model_z_trans <- readRDS(here::here("output/model/", "mcmcglmm_model_z_trans"))
+mcmcglmm_model_z_trans_nophylo <- readRDS(here::here("output/model/", "mcmcglmm_model_z_trans_nophylo"))
 }

### 2023-09-19 ms/ms.qmd
Cleaning doc

@@
 crossref:
-  fig-title: 'Figure'
+  fig-title: '**Figure**'
   fig-labels: arabic
@@
 
-```
-
-```{r, tree}
+##################################
+# Loading tree and data
+##################################
 # Bring in the fly tree
@@
 
-# Plot the tree
-  plot(tree, cex = 0.2)
+# Grab our data
+   data <- read.csv(here::here("data", "drosi_data.csv"))
+```
+
+```{r, treesetup}
+#| label: treesetup
+#| echo: false
+#| message: false
+#| include: false
 
@@
   tree2 <- force.ultrametric(tree, method = "extend")
-
-# Grab out data
-   data <- read.csv(here::here("data", "drosi_data.csv"))
 
@@
 ```{r, models}
-
+#| label: models
+#| echo: false
+#| include: false
+#| 
 ### MCMCglmm Bayesian models. Deals with missing data using data augmentation
@@
 ```{r, posteriors}
-
+#| label: posteriors
+#| echo: false
+#| include: false
+#| 
 # Extract posterior estimates
@@
 
+```{r, fig-tree}
+#| label: fig-tree
+#| fig-cap: "Phylogenetic relationships among `r length(tree$tip.label)` *Drosophila* species."
+
+# Plot the tree
+  plot(tree, cex = 0.2)
+```
+
 ```{r, posteriorsz}

### 2023-09-20 R/func.R
Added new tdtd curve calculations

@@ -19,3 +19,27 @@ tree_checks <- function(data, tree, dataCol, type = c("checks", "prune")){
     return(ape::drop.tip(tree, species_list1))
   }
 }
+
+
+#' @title TDT Curve Calculation
+#' @description This function calculates the slope and CTmax from a TDT curve fit using a linear regression
+#' @param data A dataframe containing the data  
+#' @return A data frame containing the slope and CTmax
+#' @export 
+tdt_calc <- function(data){
+
+  # Fit a linear regression to the data
+      fit <- lm(log10(t_coma) ~ assay_temp, data = data)
+  
+  # Extract the slope and CTmax
+   slope <- -1/fit$coefficients[2]
+   CTmax <- -coef(fit)[1] / coef(fit)[2]
+  
+  # Return the slope and CTmax
+    return(data.frame(z = slope, CTmax = CTmax))
+
+}
+
+#' @title TDT Curve plot
+#' 
+#' 

### 2023-09-20 ms/ms.qmd
Updates qmd to have the code to generate each new species TDTD data

@@
 ##################################
-    pacman::p_load(tidyverse,  metafor, flextable,  patchwork, ggtree, ape, phytools, viridis, MCMCglmm, brms, latex2exp, orchaRd, multcomp, bayesplot, here)
+    pacman::p_load(tidyverse,  metafor, flextable,  patchwork, ggtree, ape, phytools, viridis, MCMCglmm, brms, latex2exp, orchaRd, multcomp, bayesplot, here, readxl)
+
+##################################
+# Functions
+##################################
+
+  source(here::here("R", "func.R"))
 
@@
    data <- read.csv(here::here("data", "drosi_data.csv"))
+```
+
+```{r, newspp}
+#| label: newspp
+#| echo: false
+#| message: false
+#| include: false
+
+# Load the data
+  spp <- read_excel(here::here("data", "tdt_data.xlsx"), sheet = "data")
+
+# Calcluate the TDT
+  tdt <- split(spp, spp$group)
+
+# Calculations
+  txt <- plyr::ldply(lapply(tdt, function(x) tdt_calc(data = x)))
+
+```
+
+We can have a look at each new species' TDT curves collected by Vanessa (@fig-tdt). We can also use the data to calculate the TDT curve (slope) and $CT_{max}$ for each species (@tbl-1).
+
+```{r, fig-tdt}
+#| label: fig-tdt
+#| fig-cap: TDT curves for each species. The x-axis is temperature and the y-axis is the log10 of the time to coma. 
+#| message: false
+#| echo: false
+
+# Lets have a look at the data
+  #ggplot2::ggplot(data, ggplot2::aes(x = assay_temp, y = log10(t_coma), colour = group)) + ggplot2::geom_point() + #ggplot2::facet_wrap(~group, nrow = 2, ncol = 4) +ggplot2::labs(x = "Temperature", y = "log 10 Coma Time") + #ggplot2::theme_bw()
+
+# Fecet wrap doesn't seem to want to work
+  par(mfrow=c(2,4))
+  invisible(lapply(tdt, function(x) plot(log10(x$t_coma) ~ x$assay_temp,  main= unique(x$group), xlab = "Temperature", ylab = "log 10 Coma Time", pch = 16, col = "black")))
+```
+
+```{r, tbl-1}
+#| label: tbl-1
+#| tbl-cap: Estimated parameters for each of the 8 new species. 
+flextable(txt) %>% align(align = "center", part = "all") %>% font(fontname="Times", part = "body") %>% 
+    font(fontname="Times", part = "header")  %>%  autofit()
+```
+
+```{r, fig-tdt2}
 ```

### 2023-09-20 ms/ms.qmd
clean table

@@
 #| tbl-cap: Estimated parameters for each of the 8 new species. 
-flextable(txt) %>% align(align = "center", part = "all") %>% font(fontname="Times", part = "body") %>% 
-    font(fontname="Times", part = "header")  %>%  autofit()
+labels <- c("Species", "Slope", "CTMax")
+
+flextable(txt) %>% 
+    align(align = "center", part = "all") %>% 
+    font(fontname="Times", part = "body") %>% 
+    font(fontname="Times", part = "header")  %>%  
+    flextable::compose(i = 1, j = 1:3, part = "header", value= as_paragraph(as_b(labels))) %>% 
+    autofit()  
 ```

### 2023-09-20 ms/ms.qmd
Updated data frame
Fixed MS to use new data
Writing Box 1

@@
 # Load the data
-  spp <- read_excel(here::here("data", "tdt_data.xlsx"), sheet = "data")
+  spp <- read_csv(here::here("data", "tdt_curves.csv"))
 
@@
 
+## Box 1 – Thermal Death Time framework case studies
+
+Parameters from Thermal Death Time (TDT) curves provide a powerful tool for understanding the thermal sensitivity of organisms.
+Comparative approaches using such data can provide a way for us to evaluate differences in population and species sensitivity, and identify the key ecological and life-history traits that increase vulnerability to heat stress.
+Nonetheless, while static measures of thermal tolerance (e.g., $CT_{max}$) are widespread, parameters from TDT curves  are much less frequently estimated.
+In addition, we have a poor understanding of how dynamic (TDT) and static ($CT_{max}$) lethal measures correlate with each other and with sub-lethal measures of damage, such as critical fertility limits (CFLs).
+Identifying the correlation between lethal and sub-lethal limits can provide powerful ways to further clarify species vulnerability to heat stress that might lead to population extinction well before mortality is identified.
+
+A number of significance challenges imped are ability to take the TDT and sub-lethal frameworks into a comparative context.
+First, ethical, conservation, and logistical challenges limit our ability to measure and estimate key parameters for many species, creating significant gaps in our knowledge across the tree of life.
+Second, while there has been a focus on species-level responses, population-level information is seldom incoperated into comparative frameworks.
+
+Powerful multivariate hierachical models (i.e., multi-response models) are capable of dealing with missing data, and provide a means by which multiple physiological thermal tolerance measures can be simultaneously modelled.
+When used in conjunction with phylogenetic information across species and/or populations, such modelling approaches can provide informed estimates for species missing data and explicitly estimate the covariance between tolerance measures while accounting for their uncertainty.
+Such insight that can be used to inform policy and conservation-related decision making.
+
+Using a well studied taxonomic group, the *Drosophila* clade, we demonstrate how such an analytical approach can inform on these significant questions.
+
 ## References

### 2023-09-21 ms/ms.qmd
More writing

@@
 
+## Estimate TDT Curve Parameters for *Drosophilia*
 ```{r, newspp}
@@
 
-```{r, fig-tdt2}
-```
+## Applying Multivariate Modelling Approaches
 
@@
 First, ethical, conservation, and logistical challenges limit our ability to measure and estimate key parameters for many species, creating significant gaps in our knowledge across the tree of life.
-Second, while there has been a focus on species-level responses, population-level information is seldom incoperated into comparative frameworks.
+Second, while there has been a focus on species-level responses, population-level information is seldom incoperated into comparative frameworks limiting our ability to decompose sources of variance.
 
-Powerful multivariate hierachical models (i.e., multi-response models) are capable of dealing with missing data, and provide a means by which multiple physiological thermal tolerance measures can be simultaneously modelled.
-When used in conjunction with phylogenetic information across species and/or populations, such modelling approaches can provide informed estimates for species missing data and explicitly estimate the covariance between tolerance measures while accounting for their uncertainty.
-Such insight that can be used to inform policy and conservation-related decision making.
+Powerful multivariate hierachical models (i.e., multi-response models) are capable of dealing with missing data, and provide a means by which multiple physiological thermal tolerance measures can be modelled simultaneously [@Noble2021; @Pottier2023].
+When used in conjunction with an understanding of the phylogenetic relationships among species and/or populations, such modelling approaches can provide informed estimates for species missing data and explicitly estimate the covariance between tolerance measures while accounting for their uncertainty.
+Such insight that can be used to inform policy and conservation-related decision making for species with little to no data.
 
-Using a well studied taxonomic group, the *Drosophila* clade, we demonstrate how such an analytical approach can inform on these significant questions.
+#### Application of Multivariate Comparative Approaches using Thermal Tolerance Measures: A case study using *Drosophilids*
+
+We integrated static lethal limits ($CT_{max}$) collected by @Heerwaarden2021 and XX with dynamic lethal limit measures from TDT curces [i.e., slope ($z^{TDT}$) and intercept ($CT_{max}^{TDT}$)] estimated by XX.
+We also collected data to estimate TDT curve parameters for eight additonal species following protocols outlined in @Orsted2022.
+Critical fertility limits for *Drosophilids* were compiled from @Parratt2021 and @Heerwaarden2021.
+These data were integrated with data on lethal limits at the species-level.
+When the same species were sampled multiple times we retained within species variation in models.
+We then fit the following model in *MCMCglmm* [@]:
+
+$$
+\begin{bmatrix}
+ CT_{max} \\
+ CT_{max}^{TDT} \\
+ z^{TDT} \\
+ CFL
+\end{bmatrix} = u_{i,k} + p_{i,k} + sp_{i,k} + e_{i,k}
+$$
 

### 2023-09-21 ms/ms.qmd
Udpated doc

@@
 
-A number of significance challenges imped are ability to take the TDT and sub-lethal frameworks into a comparative context.
+A number of significant challenges imped are ability to take the TDT and sub-lethal frameworks into a comparative context.
 First, ethical, conservation, and logistical challenges limit our ability to measure and estimate key parameters for many species, creating significant gaps in our knowledge across the tree of life.
-Second, while there has been a focus on species-level responses, population-level information is seldom incoperated into comparative frameworks limiting our ability to decompose sources of variance.
+Second, while there has been a focus on species-level responses, population-level information is seldom incoperated into comparative frameworks limiting our ability to decompose sources of variance [@Pottier2023].
 
@@
 When the same species were sampled multiple times we retained within species variation in models.
-We then fit the following model in *MCMCglmm* [@]:
+We then fit the following model in R (vers.
+`r paste0(R.Version()$major, ".", R.Version()$minor)`)[@R2023] with the package *MCMCglmm* (vers.
+`r utils::packageVersion("MCMCglmm")`) [@Hadfield2010]:
 
@@
  z^{TDT} \\
- CFL
+ CFLs
 \end{bmatrix} = u_{i,k} + p_{i,k} + sp_{i,k} + e_{i,k}

### 2023-09-21 R/func.R
New functioons

@@ -40,6 +40,43 @@ tdt_calc <- function(data){
 
 }
 
-#' @title TDT Curve plot
-#' 
-#' 
+#' @title extract_blups
+#' @description This function extracts the BLUPs from a mixed model
+#' @param sol A matrix containing the BLUPs
+#' @param trait A string. The trait to be analysed. 
+#' @return A list containing the phylogenetic BLUPs and the species BLUPs
+extract_blups <- function(sol, trait){
+  animal <- paste0("trait", trait, ".animal")
+  species <- paste0("trait", trait, ".species")
+
+  phy <- sol[, grep(animal, colnames(sol))]
+  phy <- phy[,-grep("Node", colnames(phy))]
+
+  spp <- sol[, grep(species, colnames(sol))]
+
+  return(list(phy = phy, spp = spp))
+
+}
+  
+
+#' @title predict_accuracy
+#' @description This function calculates the accuracy of the model predictions compared to the data where we know the actual values
+#' @param data A dataframe containing the data
+#' @param blups A list containing the phylogenetic BLUPs and the species BLUPs
+#' @param trait The trait to be analysed. Note that this is not a string, but the name of the trait
+#' @return A data frame containing the actual and predicted values for a given trait
+predict_accuracy <- function(data, blups, trait){
+  # Get actual data
+  known <- data %>% filter(!is.na({{trait}})) %>% dplyr::select(species, {{trait}}) %>% distinct()
+  
+  # Get estimated data
+    est <- apply(blups, 2, function(x) median(x))
+    est <- est[which(colnames(blups) %in% known$species)]
+    est <- data.frame(est)
+    est$species <- row.names(est)
+
+  # Join the two
+  check <- left_join(known, est, by = "species")
+
+  return(check)
+}

### 2023-09-21 ms/ms.qmd
Updated documents

@@
 
+```{r, fig-tree}
+#| label: fig-tree
+#| fig-cap: "Phylogenetic relationships among `r length(tree$tip.label)` *Drosophila* species."
+
+# Plot the tree
+  plot(tree, cex = 0.2, type = "radial")
+```
+
 ```{r, models}
@@
     prior = list(R = list(V = diag(3), nu=1.002), 
-               G = list(G1=list(V = diag(3), nu = 1.002)))
+               G = list(G1=list(V = diag(3), nu = 1.002),
+                        G2=list(V = diag(3), nu = 1.002)))
 
@@
 
-  mcmcglmm_model <- MCMCglmm::MCMCglmm(cbind(ctmax, cfts, tdt_z) ~ -1 + trait, random = ~us(trait):animal, rcov = ~us(trait):units ,pedigree=tree3,burnin=1000, nitt = 300000, thin = 5, family = rep("gaussian", 3), prior = prior, pl = TRUE, pr = TRUE, data = data)
+  mcmcglmm_model <- MCMCglmm::MCMCglmm(cbind(ctmax, cfts, tdt_z) ~ -1 + trait, random = ~us(trait):animal + us(trait):species, rcov = ~us(trait):units ,pedigree=tree3,burnin=1000, nitt = 300000, thin = 5, family = rep("gaussian", 3), prior = prior, pl = TRUE, pr = TRUE, data = data)
 
@@
 
-  mcmcglmm_model_z_trans <- MCMCglmm::MCMCglmm(cbind(z_ctmax, z_cfts, z_tdt_z) ~ -1 + trait, random = ~us(trait):animal, rcov = ~us(trait):units, pedigree=tree3,burnin=1000, nitt = 300000, thin = 5, family = rep("gaussian", 3), prior = prior, pl = TRUE, pr = TRUE, data = data)
+  mcmcglmm_model_z_trans <- MCMCglmm::MCMCglmm(cbind(z_ctmax, z_cfts, z_tdt_z) ~ -1 + trait, random = ~us(trait):animal + us(trait):species, rcov = ~us(trait):units, pedigree=tree3,burnin=1000, nitt = 300000, thin = 5, family = rep("gaussian", 3), prior = prior, pl = TRUE, pr = TRUE, data = data)
 
   # Ditch phylogeny to see if there is support for including it or not. DIC is WAY lower for phylogenetic model, so good to include it.
-  prior2 = list(R = list(V = diag(3), nu=1.002))
-  mcmcglmm_model_z_trans_nophylo <- MCMCglmm::MCMCglmm(cbind(z_ctmax, z_cfts, z_tdt_z) ~ -1 + trait, rcov = ~us(trait):units, pedigree=tree3,burnin=1000, nitt = 300000, thin = 5, family = rep("gaussian", 3), prior = prior2, pl = TRUE, pr = TRUE, data = data)
+  prior2 = list(R = list(V = diag(3), nu=1.002), 
+               G = list(G1=list(V = diag(3), nu = 1.002)))
+  mcmcglmm_model_z_trans_nophylo <- MCMCglmm::MCMCglmm(cbind(z_ctmax, z_cfts, z_tdt_z) ~ -1 + trait, random = ~us(trait):species, rcov = ~us(trait):units, pedigree=tree3,burnin=1000, nitt = 300000, thin = 5, family = rep("gaussian", 3), prior = prior2, pl = TRUE, pr = TRUE, data = data)
 
@@
 # Extract the estimated / imputed Z for each species, but drop internal ancestral nodes
-  z_blups <- sol[, grep("traittdt_z.animal", colnames(sol))]
-  z_blups <- z_blups[,-grep("Node", colnames(z_blups))]
-  head(z_blups)
+    z_BLUPS <- extract_blups(sol, trait ="tdt_z")
+ctmax_blups <- extract_blups(sol, trait ="ctmax")
+  cfts_blups <- extract_blups(sol, trait ="cfts")
 
 # Remember that these are BLUPs. They are deviations from the intercept. So, to get back to the z we need to add the intercept to each of them.
-  z <- z_blups + sol[, "traittdt_z"]
+      z <- z_BLUPS[["phy"]] + z_BLUPS[["spp"]] + sol[, "traittdt_z"]
+  ctmax <- ctmax_blups[["phy"]] + ctmax_blups[["spp"]]+ sol[, "traitctmax"]
+   cfts <- cfts_blups[["phy"]] + cfts_blups[["spp"]]+ sol[, "traitcfts"]
 
@@
       colnames(z) <- gsub("traittdt_z.animal.", "", colnames(z))
+  colnames(ctmax) <- gsub("traitctmax.animal.", "", colnames(ctmax))
+   colnames(cfts) <- gsub("traitcfts.animal.", "", colnames(cfts))
 
 # Have a look at how well the model gets the actual data correct.
-  z_known <- data %>% filter(!is.na(tdt_z)) %>% dplyr::select(species, tdt_z) %>% distinct()
-    z_est <- apply(z, 2, function(x) median(x))
-    z_est <- z_est[which(colnames(z) %in% z_known$species)]
-    z_est <- data.frame(z_est)
-    z_est$species <- row.names(z_est)
-  check_z <- left_join(z_known, z_est, by = "species")
+    check_z <- predict_accuracy(data, z, trait = tdt_z)
+check_ctmax <- predict_accuracy(data, ctmax, trait = ctmax)
+ check_cfls <- predict_accuracy(data, cfts, trait = cfts) # Something wrong here.
+
+  # Plot the data
+    plot(check_z$tdt_z ~ check_z$est, ylab = "True Slope (z)", xlab = "Estimated Slope", pch = 16, cex = 2)
+    abline(a = 0, b = 1, col = "red")
+    plot(check_ctmax$ctmax ~ check_ctmax$est, ylab = "True CTmax", xlab = "Estimated CTmax", pch = 16, cex = 2)
+    abline(a = 0, b = 1, col = "red")
+    plot(check_cfls$cfts ~ check_cfls$est, ylab = "True CFLs", xlab = "Estimated CFLs", pch = 16, cex = 2)
+    abline(a = 0, b = 1, col = "red")
 
@@
       phylo_ctmax <- vcv[,"traitctmax:traitctmax.animal"]
+        spp_ctmax <- vcv[,"traitctmax:traitctmax.species"]
       units_ctmax <- vcv[,"traitctmax:traitctmax.units"]
 
-      h2_ctmax = phylo_ctmax / (phylo_ctmax + units_ctmax)
+      h2_ctmax = phylo_ctmax / (phylo_ctmax + spp_ctmax + units_ctmax)
       posterior.mode(h2_ctmax)
@@
       phylo_cfls <- vcv[,"traitcfts:traitcfts.animal"]
+        spp_cfls <- vcv[,"traitcfts:traitcfts.species"]
       units_cfls <- vcv[,"traitcfts:traitcfts.units"]
 
-      h2_cfls = phylo_cfls / (phylo_cfls + units_cfls)
+      h2_cfls = phylo_cfls / (phylo_cfls + spp_cfls + units_cfls)
       posterior.mode(h2_cfls)
@@
       phylo_z <- vcv[,"traittdt_z:traittdt_z.animal"]
+        spp_z <- vcv[,"traittdt_z:traittdt_z.species"]
       units_z <- vcv[,"traittdt_z:traittdt_z.units"]
 
-      h2_z = phylo_z / (phylo_z + units_z)
+      h2_z = phylo_z / (phylo_z + spp_z + units_z)
       posterior.mode(h2_z)
       HPDinterval(h2_z)
-```
-
-```{r, fig-tree}
-#| label: fig-tree
-#| fig-cap: "Phylogenetic relationships among `r length(tree$tip.label)` *Drosophila* species."
-
-# Plot the tree
-  plot(tree, cex = 0.2)
 ```
@@
 When used in conjunction with an understanding of the phylogenetic relationships among species and/or populations, such modelling approaches can provide informed estimates for species missing data and explicitly estimate the covariance between tolerance measures while accounting for their uncertainty.
-Such insight that can be used to inform policy and conservation-related decision making for species with little to no data.
+Such insight can be used to inform policy and conservation-related decision making for species with little to no data.
 
@@
 `r paste0(R.Version()$major, ".", R.Version()$minor)`)[@R2023] with the package *MCMCglmm* (vers.
-`r utils::packageVersion("MCMCglmm")`) [@Hadfield2010]:
+`r utils::packageVersion("MCMCglmm")`), which uses data augmentation during model fittng to estimate missing data [@Hadfield2010]:
 
@@
  CT_{max} \\
- CT_{max}^{TDT} \\
  z^{TDT} \\
@@
 
+Models including phylogeny were better supported than models without phylogeny ($\Delta DIC$ = `r mcmcglmm_model_z_trans_nophylo$DIC - mcmcglmm_model_z_trans$DIC`), with all traits exhibiting a moderate to strong phylogenetic signal (Phylogenetic $H^2$: $CT_{max}$ – `r posterior.mode(h2_ctmax)`, 95% CI: `r HPDinterval(h2_ctmax)[,"lower"]` to `r HPDinterval(h2_ctmax)[,"upper"]`; $z^{TDT}$ – `r posterior.mode(h2_z)`, 95% CI: `r HPDinterval(h2_z)[,"lower"]` to `r HPDinterval(h2_z)[,"upper"]`; CFLs – `r posterior.mode(h2_cfls)`, 95% CI: `r HPDinterval(h2_cfls)[,"lower"]` to `r HPDinterval(h2_cfls)[,"upper"]`).
+The model predicted known data reasonably well for most traits (correlation between actual and estimated data: $z^{TDT}$ – `r cor.test(check_z$tdt_z, check_z$est)$estimate`, 95% CI: `r cor.test(check_z$tdt_z, check_z$est)$conf.int[1]` to `r cor.test(check_z$tdt_z, check_z$est)$conf.int[2]`, p < 0.001).
+
+```{r, fig-treesub}
+#| label: fig-treesub
+#| fig-cap: Species included in the dataset and their phylogenetic relationships.
+
+# prune tree
+t <- tree_checks(data, tree, dataCol = "species", type = "prune")
+
+layout.matrix <- matrix(c(1, 1, 1, 2,3,4), nrow = 3, ncol = 2) # Phylogeny takes up 3 panels (plot = 1) and the others take up a single panel.
+
+layout(mat = layout.matrix) 
+# Plot 1
+plot(t, type = "tidy", cex = 0.65)
+
+# Plot 2
+plot(check_z$tdt_z ~ check_z$est, ylab = "TDT Slope (z)", xlab = "Estimated TDT Slope", pch = 16, cex = 3, cex.lab = 2)
+abline(a = 0, b = 1, col = "red")
+
+```
+
 ## References

### 2023-09-21 R/func.R
Fixed bug in function

@@ -66,8 +66,9 @@ extract_blups <- function(sol, trait){
 #' @param trait The trait to be analysed. Note that this is not a string, but the name of the trait
 #' @return A data frame containing the actual and predicted values for a given trait
 predict_accuracy <- function(data, blups, trait){
-  # Get actual data
-  known <- data %>% filter(!is.na({{trait}})) %>% dplyr::select(species, {{trait}}) %>% distinct()
+  # Get actual data. Note that we have species replicated multiple times. Stick to among species level so need to average the replicate populations of a given species. BLUPs are only estimated for a single species. 
+  
+  known <- data %>% filter(!is.na({{trait}})) %>% dplyr::select(species, {{trait}}) %>% group_by(species) %>% summarise(trait = mean({{trait}}))
   
   # Get estimated data
     est <- apply(blups, 2, function(x) median(x))

### 2023-09-21 R/func.R
Function to calculate the phylogenetic correlation between two traits

@@ -67,7 +67,7 @@ extract_blups <- function(sol, trait){
 #' @return A data frame containing the actual and predicted values for a given trait
 predict_accuracy <- function(data, blups, trait){
   # Get actual data. Note that we have species replicated multiple times. Stick to among species level so need to average the replicate populations of a given species. BLUPs are only estimated for a single species. 
-  
+
   known <- data %>% filter(!is.na({{trait}})) %>% dplyr::select(species, {{trait}}) %>% group_by(species) %>% summarise(trait = mean({{trait}}))
   
   # Get estimated data
@@ -80,4 +80,29 @@ predict_accuracy <- function(data, blups, trait){
   check <- left_join(known, est, by = "species")
 
   return(check)
+}
+
+#' @title phylo_corrs
+#' @description This function calculates the phylogenetic correlation between two traits
+#' @param vcv A variance covariance matrix
+#' @param trait1 The first trait to be analysed
+#' @param trait2 The second trait to be analysed
+#' @return A vector containing the posterior distribution of the phylogenetic correlation between the two traits
+phylo_corrs <- function(vcv, trait1, trait2){
+    
+    # Grep strings
+    animalt1 <- paste0("trait", trait1)
+    animalt2 <- paste0("trait", trait2)
+    
+    # Extract the variance components
+    vart1 <- vcv[,grep(paste0(animalt1, ":", animalt1, ".animal"), colnames(vcv))]
+    vart2 <- vcv[,grep(paste0(animalt2, ":", animalt2, ".animal"), colnames(vcv))]
+
+    # extract covariance between traits
+    cov_name <- paste0(animalt1, ":", animalt2, ".animal")
+
+    # Calculate correlatoon
+    r_phylo <- vcv[,cov_name] / sqrt(vart1 * vart2)
+
+    return(r_phylo)
 }

### 2023-09-21 ms/ms.qmd
Finished first draft of Box1. Need to make a pretty figure

@@
   # Plot the data
-    plot(check_z$tdt_z ~ check_z$est, ylab = "True Slope (z)", xlab = "Estimated Slope", pch = 16, cex = 2)
+    plot(check_z$trait ~ check_z$est, ylab = "True Slope (z)", xlab = "Estimated Slope", pch = 16, cex = 2)
     abline(a = 0, b = 1, col = "red")
-    plot(check_ctmax$ctmax ~ check_ctmax$est, ylab = "True CTmax", xlab = "Estimated CTmax", pch = 16, cex = 2)
+    plot(check_ctmax$trait ~ check_ctmax$est, ylab = "True CTmax", xlab = "Estimated CTmax", pch = 16, cex = 2)
     abline(a = 0, b = 1, col = "red")
-    plot(check_cfls$cfts ~ check_cfls$est, ylab = "True CFLs", xlab = "Estimated CFLs", pch = 16, cex = 2)
+    plot(check_cfls$trait ~ check_cfls$est, ylab = "True CFLs", xlab = "Estimated CFLs", pch = 16, cex = 2)
     abline(a = 0, b = 1, col = "red")
@@
       h2_ctmax = phylo_ctmax / (phylo_ctmax + spp_ctmax + units_ctmax)
+      spp_ctmax <- spp_ctmax / (phylo_ctmax + spp_ctmax + units_ctmax)
       posterior.mode(h2_ctmax)
       HPDinterval(h2_ctmax)
+
+      posterior.mode(spp_ctmax)
+      HPDinterval(spp_ctmax)
 
@@
       h2_cfls = phylo_cfls / (phylo_cfls + spp_cfls + units_cfls)
+     spp_cfls = spp_cfls / (phylo_cfls + spp_cfls + units_cfls)
       posterior.mode(h2_cfls)
       HPDinterval(h2_cfls)
+
+       posterior.mode(spp_cfls)
+      HPDinterval(spp_cfls)
 
@@
       h2_z = phylo_z / (phylo_z + spp_z + units_z)
+      spp_z = spp_z / (phylo_z + spp_z + units_z)
       posterior.mode(h2_z)
       HPDinterval(h2_z)
+
+      posterior.mode(spp_z)
+      HPDinterval(spp_z)
 ```
 
+```{r, correlations}
+
+# Estimate phylogenetic correlatons between traits. 
+    # Trait variances 
+   phylocor_ctmax_tdt <- phylo_corrs(vcv, trait1 = "tdt_z", trait2 = "ctmax")
+   posterior.mode(phylocor_ctmax_tdt)
+   HPDinterval(phylocor_ctmax_tdt)
+
+   phylocor_ctmax_cfls <- phylo_corrs(vcv, trait1 = "cfts", trait2 = "ctmax")
+   posterior.mode(phylocor_ctmax_cfls)
+   HPDinterval(phylocor_ctmax_cfls)
+
+  phylocor_tdt_cfls <- phylo_corrs(vcv, trait1 = "tdt_z", trait2 = "cfts")
+   posterior.mode(phylocor_tdt_cfls)
+   HPDinterval(phylocor_tdt_cfls)
+
+```
 ```{r, posteriorsz}
@@
 
-Models including phylogeny were better supported than models without phylogeny ($\Delta DIC$ = `r mcmcglmm_model_z_trans_nophylo$DIC - mcmcglmm_model_z_trans$DIC`), with all traits exhibiting a moderate to strong phylogenetic signal (Phylogenetic $H^2$: $CT_{max}$ – `r posterior.mode(h2_ctmax)`, 95% CI: `r HPDinterval(h2_ctmax)[,"lower"]` to `r HPDinterval(h2_ctmax)[,"upper"]`; $z^{TDT}$ – `r posterior.mode(h2_z)`, 95% CI: `r HPDinterval(h2_z)[,"lower"]` to `r HPDinterval(h2_z)[,"upper"]`; CFLs – `r posterior.mode(h2_cfls)`, 95% CI: `r HPDinterval(h2_cfls)[,"lower"]` to `r HPDinterval(h2_cfls)[,"upper"]`).
-The model predicted known data reasonably well for most traits (correlation between actual and estimated data: $z^{TDT}$ – `r cor.test(check_z$tdt_z, check_z$est)$estimate`, 95% CI: `r cor.test(check_z$tdt_z, check_z$est)$conf.int[1]` to `r cor.test(check_z$tdt_z, check_z$est)$conf.int[2]`, p < 0.001).
+Models including phylogeny were better supported than models without phylogeny ($\Delta DIC$ = `r mcmcglmm_model_z_trans_nophylo$DIC - mcmcglmm_model_z_trans$DIC`), with all traits exhibiting a moderate to strong phylogenetic signal indicating evolutionary conservatism in heat tolerance (Phylogenetic $H^2$: $CT_{max}$ – `r posterior.mode(h2_ctmax)`, 95% CI: `r HPDinterval(h2_ctmax)[,"lower"]` to `r HPDinterval(h2_ctmax)[,"upper"]`; $z^{TDT}$ – `r posterior.mode(h2_z)`, 95% CI: `r HPDinterval(h2_z)[,"lower"]` to `r HPDinterval(h2_z)[,"upper"]`; CFLs – `r posterior.mode(h2_cfls)`, 95% CI: `r HPDinterval(h2_cfls)[,"lower"]` to `r HPDinterval(h2_cfls)[,"upper"]`).
+Nonetheless, there was within species variation ($I^2_{species}$: $CT_{max}$ – `r posterior.mode(spp_ctmax)`, 95% CI: `r HPDinterval(spp_ctmax)[,"lower"]` to `r HPDinterval(spp_ctmax)[,"upper"]`; $z^{TDT}$ – `r posterior.mode(spp_z)`, 95% CI: `r HPDinterval(spp_z)[,"lower"]` to `r HPDinterval(spp_z)[,"upper"]`; CFLs – `r posterior.mode(spp_cfls)`, 95% CI: `r HPDinterval(spp_cfls)[,"lower"]` to `r HPDinterval(spp_cfls)[,"upper"]`).
+
+The model predicted known data reasonably well for most traits (correlation between known values and values estimated from the model: $z^{TDT}$ – `r cor.test(check_z$trait, check_z$est)$estimate`, 95% CI: `r cor.test(check_z$trait, check_z$est)$conf.int[1]` to `r cor.test(check_z$trait, check_z$est)$conf.int[2]`, p < 0.001; $CT_{max}$ – `r cor.test(check_ctmax$trait, check_ctmax$est)$estimate`, 95% CI: `r cor.test(check_ctmax$trait, check_ctmax$est)$conf.int[1]` to `r cor.test(check_ctmax$trait, check_ctmax$est)$conf.int[2]`, p < 0.001; CFLs – `r cor.test(check_cfls$trait, check_cfls$est)$estimate`, 95% CI: `r cor.test(check_cfls$trait, check_cfls$est)$conf.int[1]` to `r cor.test(check_cfls$trait, check_cfls$est)$conf.int[2]`, p < 0.001), suggesting that even this basic model could be used to provide sensible quantitative measures for *Drosophila* species without heat tolerance estimates.
+While modelling approaches such as these are not a replacement for actual empirical data, they could be used to provide informed judgements on thermal tolerance in the absence of such information, but should be tested (using cross-validation) and refined when necessary.
+
+Interestingly, there was a reasonably strong phylogenetic correlation between $CT_{max}$ and critical fertility limits in *Drosophilids* (r = `r posterior.mode(phylocor_ctmax_cfls)`, 95% CI: `r HPDinterval(phylocor_ctmax_cfls)[,"lower"]` to `r HPDinterval(phylocor_ctmax_cfls)[,"upper"]`), but not between $CT_{max}$, CFLs and the slope from TDT curves (cor($CT_{max}$, $z^{TDT}$) = `r posterior.mode(phylocor_ctmax_tdt)`, 95% CI: `r HPDinterval(phylocor_ctmax_tdt)[,"lower"]` to `r HPDinterval(phylocor_ctmax_tdt)[,"upper"]`; cor(CFLs, $z^{TDT}$) = `r posterior.mode(phylocor_tdt_cfls)`, 95% CI: `r HPDinterval(phylocor_tdt_cfls)[,"lower"]` to `r HPDinterval(phylocor_tdt_cfls)[,"upper"]`).
+These results suggest that sub-lethal measures, such as CFLs strongly coincide with a species' upper lethal limit.
+It would be interesting to understand if this is a general phenomena across more taxonomic groups.
 
@@
 # Plot 2
-plot(check_z$tdt_z ~ check_z$est, ylab = "TDT Slope (z)", xlab = "Estimated TDT Slope", pch = 16, cex = 3, cex.lab = 2)
+plot(check_z$trait ~ check_z$est, ylab = "TDT Slope (z)", xlab = "Estimated TDT Slope", pch = 16, cex = 3, cex.lab = 2)
 abline(a = 0, b = 1, col = "red")

### 2023-09-21 ms/ms.qmd
Better subtitles to grab reader

@@
 
+##### Evolutionary signatures of thermal tolerance in *Drosophilids*
+
 Models including phylogeny were better supported than models without phylogeny ($\Delta DIC$ = `r mcmcglmm_model_z_trans_nophylo$DIC - mcmcglmm_model_z_trans$DIC`), with all traits exhibiting a moderate to strong phylogenetic signal indicating evolutionary conservatism in heat tolerance (Phylogenetic $H^2$: $CT_{max}$ – `r posterior.mode(h2_ctmax)`, 95% CI: `r HPDinterval(h2_ctmax)[,"lower"]` to `r HPDinterval(h2_ctmax)[,"upper"]`; $z^{TDT}$ – `r posterior.mode(h2_z)`, 95% CI: `r HPDinterval(h2_z)[,"lower"]` to `r HPDinterval(h2_z)[,"upper"]`; CFLs – `r posterior.mode(h2_cfls)`, 95% CI: `r HPDinterval(h2_cfls)[,"lower"]` to `r HPDinterval(h2_cfls)[,"upper"]`).
@@
 
+##### Evolutionary relationships can help understand thermal tolerance in species lacking thermal tolerance data
 The model predicted known data reasonably well for most traits (correlation between known values and values estimated from the model: $z^{TDT}$ – `r cor.test(check_z$trait, check_z$est)$estimate`, 95% CI: `r cor.test(check_z$trait, check_z$est)$conf.int[1]` to `r cor.test(check_z$trait, check_z$est)$conf.int[2]`, p < 0.001; $CT_{max}$ – `r cor.test(check_ctmax$trait, check_ctmax$est)$estimate`, 95% CI: `r cor.test(check_ctmax$trait, check_ctmax$est)$conf.int[1]` to `r cor.test(check_ctmax$trait, check_ctmax$est)$conf.int[2]`, p < 0.001; CFLs – `r cor.test(check_cfls$trait, check_cfls$est)$estimate`, 95% CI: `r cor.test(check_cfls$trait, check_cfls$est)$conf.int[1]` to `r cor.test(check_cfls$trait, check_cfls$est)$conf.int[2]`, p < 0.001), suggesting that even this basic model could be used to provide sensible quantitative measures for *Drosophila* species without heat tolerance estimates.
@@
 
+##### Static lethal limits are evolutionarily correlated with sub-lethal limits in *Drosophilids*
 Interestingly, there was a reasonably strong phylogenetic correlation between $CT_{max}$ and critical fertility limits in *Drosophilids* (r = `r posterior.mode(phylocor_ctmax_cfls)`, 95% CI: `r HPDinterval(phylocor_ctmax_cfls)[,"lower"]` to `r HPDinterval(phylocor_ctmax_cfls)[,"upper"]`), but not between $CT_{max}$, CFLs and the slope from TDT curves (cor($CT_{max}$, $z^{TDT}$) = `r posterior.mode(phylocor_ctmax_tdt)`, 95% CI: `r HPDinterval(phylocor_ctmax_tdt)[,"lower"]` to `r HPDinterval(phylocor_ctmax_tdt)[,"upper"]`; cor(CFLs, $z^{TDT}$) = `r posterior.mode(phylocor_tdt_cfls)`, 95% CI: `r HPDinterval(phylocor_tdt_cfls)[,"lower"]` to `r HPDinterval(phylocor_tdt_cfls)[,"upper"]`).

### 2023-09-21 ms/ms.qmd
Updted docs. Still need to fix figure diemnsions

@@
 ##################################
-    pacman::p_load(tidyverse,  metafor, flextable,  patchwork, ggtree, ape, phytools, viridis, MCMCglmm, brms, latex2exp, orchaRd, multcomp, bayesplot, here, readxl)
+    pacman::p_load(tidyverse,  metafor, flextable,  patchwork, ggtree, ape, phytools, viridis, MCMCglmm, brms, latex2exp, orchaRd, multcomp, bayesplot, here, readxl, magick)
 
@@
 
-##### Evolutionary relationships can help understand thermal tolerance in species lacking thermal tolerance data
+##### Evolutionary relationships can help understand thermal tolerance in species lacking data
 The model predicted known data reasonably well for most traits (correlation between known values and values estimated from the model: $z^{TDT}$ – `r cor.test(check_z$trait, check_z$est)$estimate`, 95% CI: `r cor.test(check_z$trait, check_z$est)$conf.int[1]` to `r cor.test(check_z$trait, check_z$est)$conf.int[2]`, p < 0.001; $CT_{max}$ – `r cor.test(check_ctmax$trait, check_ctmax$est)$estimate`, 95% CI: `r cor.test(check_ctmax$trait, check_ctmax$est)$conf.int[1]` to `r cor.test(check_ctmax$trait, check_ctmax$est)$conf.int[2]`, p < 0.001; CFLs – `r cor.test(check_cfls$trait, check_cfls$est)$estimate`, 95% CI: `r cor.test(check_cfls$trait, check_cfls$est)$conf.int[1]` to `r cor.test(check_cfls$trait, check_cfls$est)$conf.int[2]`, p < 0.001), suggesting that even this basic model could be used to provide sensible quantitative measures for *Drosophila* species without heat tolerance estimates.
@@
 #| fig-cap: Species included in the dataset and their phylogenetic relationships.
+#| echo: false
+#| message: false
 
@@
 
+rerun = FALSE
+if(rerun){
+  pdf(height = 10.2, width = 8.1, file = here::here("output", "fig", "fig1.pdf"))
   layout(mat = layout.matrix) 
@@
   # Plot 2
-plot(check_z$trait ~ check_z$est, ylab = "TDT Slope (z)", xlab = "Estimated TDT Slope", pch = 16, cex = 3, cex.lab = 2)
+  plot(check_z$trait ~ check_z$est, ylab = "TDT Slope (z)", xlab = "Estimated TDT Slope", pch = 16, cex = 3, cex.lab = 1.4)
   abline(a = 0, b = 1, col = "red")
-
+  dev.off()
+} else {
+  fig <- magick::image_read(here::here("output", "fig", "fig1.pdf"))
+  plot(fig)
+}
 ```

### 2023-09-22 ms/ms.qmd
trying to fix figure

@@
 
+rerun = FALSE
+if(rerun){
+
   layout.matrix <- matrix(c(1, 1, 1, 2,3,4), nrow = 3, ncol = 2) # Phylogeny takes up 3 panels (plot = 1) and the others take up a single panel.
 
-rerun = FALSE
-if(rerun){
   pdf(height = 10.2, width = 8.1, file = here::here("output", "fig", "fig1.pdf"))

### 2023-09-23 ms/ms.qmd
Added some changes to text and reminders on what to write

@@
 
+$$
+p_{i,k} \sim N\left(0, \right) \\
+
+sp_{i,k} \sim N\left(0, \right) \\
+
+e_{i,k} \sim N\left(0,  \right) \\
+$$
+
 ##### Evolutionary signatures of thermal tolerance in *Drosophilids*
 
+We obtained data for `r` species ().
 Models including phylogeny were better supported than models without phylogeny ($\Delta DIC$ = `r mcmcglmm_model_z_trans_nophylo$DIC - mcmcglmm_model_z_trans$DIC`), with all traits exhibiting a moderate to strong phylogenetic signal indicating evolutionary conservatism in heat tolerance (Phylogenetic $H^2$: $CT_{max}$ – `r posterior.mode(h2_ctmax)`, 95% CI: `r HPDinterval(h2_ctmax)[,"lower"]` to `r HPDinterval(h2_ctmax)[,"upper"]`; $z^{TDT}$ – `r posterior.mode(h2_z)`, 95% CI: `r HPDinterval(h2_z)[,"lower"]` to `r HPDinterval(h2_z)[,"upper"]`; CFLs – `r posterior.mode(h2_cfls)`, 95% CI: `r HPDinterval(h2_cfls)[,"lower"]` to `r HPDinterval(h2_cfls)[,"upper"]`).
-Nonetheless, there was within species variation ($I^2_{species}$: $CT_{max}$ – `r posterior.mode(spp_ctmax)`, 95% CI: `r HPDinterval(spp_ctmax)[,"lower"]` to `r HPDinterval(spp_ctmax)[,"upper"]`; $z^{TDT}$ – `r posterior.mode(spp_z)`, 95% CI: `r HPDinterval(spp_z)[,"lower"]` to `r HPDinterval(spp_z)[,"upper"]`; CFLs – `r posterior.mode(spp_cfls)`, 95% CI: `r HPDinterval(spp_cfls)[,"lower"]` to `r HPDinterval(spp_cfls)[,"upper"]`).
+While the model also suggested that there was within species variation, missing data and lack of replicate species populations make conclusions premature ($I^2_{species}$: $CT_{max}$ – `r posterior.mode(spp_ctmax)`, 95% CI: `r HPDinterval(spp_ctmax)[,"lower"]` to `r HPDinterval(spp_ctmax)[,"upper"]`; $z^{TDT}$ – `r posterior.mode(spp_z)`, 95% CI: `r HPDinterval(spp_z)[,"lower"]` to `r HPDinterval(spp_z)[,"upper"]`; CFLs – `r posterior.mode(spp_cfls)`, 95% CI: `r HPDinterval(spp_cfls)[,"lower"]` to `r HPDinterval(spp_cfls)[,"upper"]`).
 

### 2023-09-24 ms/ms.qmd
Fixed equation

@@
    data <- read.csv(here::here("data", "drosi_data.csv"))
+
+##################################
+# Summary Data
+##################################
+
+sum <- data  %>% group_by(species)  %>% summarise(ctmax = mean(ctmax, na.rm = TRUE),
+                                                   cfls = mean(cfts, na.rm = TRUE),
+                                                   tdt_z = mean(cfts, na.rm = TRUE))
+
+sum2 <- sapply(sum, function(x) sum(!is.na(x)))
 ```
@@
   tree3$node.label <- NULL
+
+# prune tree
+  t <- tree_checks(data, tree, dataCol = "species", type = "prune")
 
@@
 ```
+
 ```{r, posteriorsz}
@@
 These data were integrated with data on lethal limits at the species-level.
+Overall, we obtained data for `r length(unique(t$tip.label))` species ($N_{CT_{max}}$ = `r sum2[2]` species; $N_{z^{TDT}}$ = `r sum2[4]` species and $N_{CFLs}$ = `r sum2[3]` species).
 When the same species were sampled multiple times we retained within species variation in models.
@@
 $$
-p_{i,k} \sim N\left(0, \right) \\
-
-sp_{i,k} \sim N\left(0, \right) \\
-
-e_{i,k} \sim N\left(0,  \right) \\
+\begin{aligned}
+& p_{i,k} \sim N\left(0, \sigma_{p}^2 \textbf{A} \right) \\
+& sp_{i,k} \sim N\left(0, \sigma_{sp}^2\textbf{I} \right) \\
+& e_{i,k} \sim N\left(0,  \sigma_{e}^2\textbf{I}\right) \\
+\end{aligned}
 $$
@@
 
-We obtained data for `r` species ().
  Models including phylogeny were better supported than models without phylogeny ($\Delta DIC$ = `r mcmcglmm_model_z_trans_nophylo$DIC - mcmcglmm_model_z_trans$DIC`), with all traits exhibiting a moderate to strong phylogenetic signal indicating evolutionary conservatism in heat tolerance (Phylogenetic $H^2$: $CT_{max}$ – `r posterior.mode(h2_ctmax)`, 95% CI: `r HPDinterval(h2_ctmax)[,"lower"]` to `r HPDinterval(h2_ctmax)[,"upper"]`; $z^{TDT}$ – `r posterior.mode(h2_z)`, 95% CI: `r HPDinterval(h2_z)[,"lower"]` to `r HPDinterval(h2_z)[,"upper"]`; CFLs – `r posterior.mode(h2_cfls)`, 95% CI: `r HPDinterval(h2_cfls)[,"lower"]` to `r HPDinterval(h2_cfls)[,"upper"]`).
@@
 #| message: false
-
-# prune tree
-t <- tree_checks(data, tree, dataCol = "species", type = "prune")
 

### 2023-09-24 ms/ms.qmd
Updated MS

@@
                                                    cfls = mean(cfts, na.rm = TRUE),
-                                                   tdt_z = mean(cfts, na.rm = TRUE))
+                                                   tdt_z = mean(cfts, na.rm = TRUE))  %>% data.frame()
 
-sum2 <- sapply(sum, function(x) sum(!is.na(x)))
+num_spp <- sapply(sum, function(x) sum(!is.na(x)))
 ```
@@
 These data were integrated with data on lethal limits at the species-level.
-Overall, we obtained data for `r length(unique(t$tip.label))` species ($N_{CT_{max}}$ = `r sum2[2]` species; $N_{z^{TDT}}$ = `r sum2[4]` species and $N_{CFLs}$ = `r sum2[3]` species).
+Overall, we obtained data for `r length(unique(t$tip.label))` species ($N_{CT_{max}}$ = `r num_spp[2]-5` species; $N_{z^{TDT}}$ = `r num_spp[4]-4` species and $N_{CFLs}$ = `r num_spp[3]-4` species).
 When the same species were sampled multiple times we retained within species variation in models.
@@
  CFLs
-\end{bmatrix} = u_{i,k} + p_{i,k} + sp_{i,k} + e_{i,k}
+\end{bmatrix} = u_{i,k} + p_{k} + sp_{k} + e_{i,k}
 $$
@@
 \begin{aligned}
-& p_{i,k} \sim N\left(0, \sigma_{p}^2 \textbf{A} \right) \\
-& sp_{i,k} \sim N\left(0, \sigma_{sp}^2\textbf{I} \right) \\
+& p_{k} \sim N\left(0, \sigma_{p}^2 \textbf{A} \right) \\
+& sp_{k} \sim N\left(0, \sigma_{sp}^2\textbf{I} \right) \\
 & e_{i,k} \sim N\left(0,  \sigma_{e}^2\textbf{I}\right) \\
@@
 $$
+
+where $p_{k}$ and $sp_{k}$ are phylogenetic (evolutionary) and population level effects (assumed to be sampled from a normal distribution, $\sim N\left(0, \sigma_{sp/p}^2 \right)$) with the evolutionary relatedness being structured according to the phylogenetic relatesness matrix ($\textbf{A}$).
 

### 2023-09-25 ms/ms.qmd
First draft of box

@@
 
-We can have a look at each new species' TDT curves collected by Vanessa (@fig-tdt). We can also use the data to calculate the TDT curve (slope) and $CT_{max}$ for each species (@tbl-1).
+We can have a look at each new species' TDT curves collected by Vanessa (@fig-tdt). We can also use the data to calculate the TDT curve (slope) and $CT_{max}$ for each species (Table 1).
 
@@
 ```{r, correlations}
-
+#| label: correlations
+#| echo: false
+#| include: false
 # Estimate phylogenetic correlatons between traits. 
@@
 These data were integrated with data on lethal limits at the species-level.
-Overall, we obtained data for `r length(unique(t$tip.label))` species ($N_{CT_{max}}$ = `r num_spp[2]-5` species; $N_{z^{TDT}}$ = `r num_spp[4]-4` species and $N_{CFLs}$ = `r num_spp[3]-4` species).
+Overall, we obtained data for `r num_spp[1]` species ($N_{CT_{max}}$ = `r num_spp[2]` species; $N_{z^{TDT}}$ = `r num_spp[4]` species and $N_{CFLs}$ = `r num_spp[3]` species).
 When the same species were sampled multiple times we retained within species variation in models.
@@
 \begin{aligned}
-& p_{k} \sim N\left(0, \sigma_{p}^2 \textbf{A} \right) \\
-& sp_{k} \sim N\left(0, \sigma_{sp}^2\textbf{I} \right) \\
-& e_{i,k} \sim N\left(0,  \sigma_{e}^2\textbf{I}\right) \\
+ & p_{k} \sim N\left(0, \sigma^2_{p} \textbf{A}\right) \\ 
+ & sp_{k} \sim N\left(0, \sigma^2_{sp} \textbf{I}\right) \\
+ & e_{i,k} \sim N\left(0, \sigma^2_{e} \textbf{I}\right)
 \end{aligned}
@@
 
-where $p_{k}$ and $sp_{k}$ are phylogenetic (evolutionary) and population level effects (assumed to be sampled from a normal distribution, $\sim N\left(0, \sigma_{sp/p}^2 \right)$) with the evolutionary relatedness being structured according to the phylogenetic relatesness matrix ($\textbf{A}$).
+where $u_{i,k}$ is the overall mean (intercept) for each trait for each observation, *i*, for species, *k*; $p_{k}$ and $sp_{k}$ are phylogenetic (evolutionary) and species effects (i.e., ecological species effects) (assumed to be sampled from a normal distribution) with the evolutionary relatedness being structured according to the phylogenetic relatedness matrix ($\textbf{A}$) as opposed to an identity matrix ($\textbf{I}$).
+Heat tolerance measures are modelled according to a multivariate normal distribution.
+Covariances between thermal tolerance measures at each random effect level are also estimated.
 

### 2023-09-27 ms/ms.qmd
Integrating Vanessa's comments. May need to edit last paragraph to reduce some duplication.

@@
 
-## Box 1 – Thermal Death Time framework case studies
+## Box 1 – Comparing the utility of TDT Curves to other measures of heat tolerance within a comparative framework
 
 Parameters from Thermal Death Time (TDT) curves provide a powerful tool for understanding the thermal sensitivity of organisms.
-Comparative approaches using such data can provide a way for us to evaluate differences in population and species sensitivity, and identify the key ecological and life-history traits that increase vulnerability to heat stress.
-Nonetheless, while static measures of thermal tolerance (e.g., $CT_{max}$) are widespread, parameters from TDT curves  are much less frequently estimated.
-In addition, we have a poor understanding of how dynamic (TDT) and static ($CT_{max}$) lethal measures correlate with each other and with sub-lethal measures of damage, such as critical fertility limits (CFLs).
-Identifying the correlation between lethal and sub-lethal limits can provide powerful ways to further clarify species vulnerability to heat stress that might lead to population extinction well before mortality is identified.
+But there are drawbacks to estimating TDT curves in comparison to other lethal (e.g.
+$CT_{max}$) and sublethal (CFL) measures.
+TDT curves are logistically challenging requiring more individuals for reliable estimates than other measures like $CT_{max}$ and CFL’s.
+TDT curves (and $CT_{max}$) also cannot be used for all organisms due to ethical and conservation considerations.
+Because of these limitations data on thermal sensitivity/heat vulnerability is missing for many species, creating significant knowledge gaps across the tree of life.
 
-A number of significant challenges imped are ability to take the TDT and sub-lethal frameworks into a comparative context.
-First, ethical, conservation, and logistical challenges limit our ability to measure and estimate key parameters for many species, creating significant gaps in our knowledge across the tree of life.
-Second, while there has been a focus on species-level responses, population-level information is seldom incoperated into comparative frameworks limiting our ability to decompose sources of variance [@Pottier2023].
+Phylogenetic comparative approaches could be leveraged to help fill these knowledge gaps because closely related species are often more likely to share similar traits owing to their evolutionary history (phylogenetic niche conservatism) (ref).
+If heat tolerance is linked to the phylogeny, shared characteristics that make species more or less heat tolerant will not be randomly distributed across the phylogeny but biased by species groups (ref).
+We can can use evolutionary relatedness to predict species thermal sensitivities and such insights can be used to inform policy and conservation related decisions making for species with little to no data.
 
-Powerful multivariate hierachical models (i.e., multi-response models) are capable of dealing with missing data, and provide a means by which multiple physiological thermal tolerance measures can be modelled simultaneously [@Noble2021; @Pottier2023].
-When used in conjunction with an understanding of the phylogenetic relationships among species and/or populations, such modelling approaches can provide informed estimates for species missing data and explicitly estimate the covariance between tolerance measures while accounting for their uncertainty.
-Such insight can be used to inform policy and conservation-related decision making for species with little to no data.
+While the comparative approach can help fill knowledge gaps it does not entirely solve the issue that TDT curves cannot be estimated for many organisms (logistical and ethical).
+$CT_{max}$ is less logistically challenging but also has ethical limitations.
+Ethical issues could be alleviated by the use of sublethal measures of thermal tolerance like CFL’s (ref), but we do not know how well lethal (TDT curves/$CT_{max}$) and sublethal (CFL’s) measures capture thermal sensitivities/vulnerabilities.
+Identifying whether a correlation exists between lethal and sub-lethal limits will help us understand the mechanisms of vulnerability to heat stress that might lead to population extinction well before mortality is identified.
+And if these traits are correlated different measures of thermal tolerance can be used for different organisms.
 
 #### Application of Multivariate Comparative Approaches using Thermal Tolerance Measures: A case study using *Drosophilids*
+
+Here we use powerful multivariate hierachical models (i.e., multi-response models) that are capable of dealing with missing data, and provide a means by which multiple physiological thermal tolerance measures can be modelled simultaneously [@Pottier2023; @Noble2021].
+When used in conjunction with an understanding of the phylogenetic relationships among species and/or populations, such modelling approaches can provide informed estimates for species missing data and explicitly estimate the covariance between tolerance measures while accounting for their uncertainty.
 

### 2023-09-27 ms/ms.qmd
Updating with Vanessa's revisions and streamlining a bit more

@@
 ---
-title: "Box 1 - TDT Curves"
+title: ""
 bibliography: ../bib/refs.bib
@@
 
-## Estimate TDT Curve Parameters for *Drosophilia*
-```{r, newspp}
-#| label: newspp
-#| echo: false
-#| message: false
-#| include: false
-
-# Load the data
-  spp <- read_csv(here::here("data", "tdt_curves.csv"))
-
-# Calcluate the TDT
-  tdt <- split(spp, spp$group)
-
-# Calculations
-  txt <- plyr::ldply(lapply(tdt, function(x) tdt_calc(data = x)))
-
-```
-
-We can have a look at each new species' TDT curves collected by Vanessa (@fig-tdt). We can also use the data to calculate the TDT curve (slope) and $CT_{max}$ for each species (Table 1).
-
-```{r, fig-tdt}
-#| label: fig-tdt
-#| fig-cap: TDT curves for each species. The x-axis is temperature and the y-axis is the log10 of the time to coma. 
-#| message: false
-#| echo: false
-
-# Lets have a look at the data
-  #ggplot2::ggplot(data, ggplot2::aes(x = assay_temp, y = log10(t_coma), colour = group)) + ggplot2::geom_point() + #ggplot2::facet_wrap(~group, nrow = 2, ncol = 4) +ggplot2::labs(x = "Temperature", y = "log 10 Coma Time") + #ggplot2::theme_bw()
-
-# Fecet wrap doesn't seem to want to work
-  par(mfrow=c(2,4))
-  invisible(lapply(tdt, function(x) plot(log10(x$t_coma) ~ x$assay_temp,  main= unique(x$group), xlab = "Temperature", ylab = "log 10 Coma Time", pch = 16, col = "black")))
-```
-
-```{r, tbl-1}
-#| label: tbl-1
-#| tbl-cap: Estimated parameters for each of the 8 new species. 
-labels <- c("Species", "Slope", "CTMax")
-
-flextable(txt) %>% 
-    align(align = "center", part = "all") %>% 
-    font(fontname="Times", part = "body") %>% 
-    font(fontname="Times", part = "header")  %>%  
-    flextable::compose(i = 1, j = 1:3, part = "header", value= as_paragraph(as_b(labels))) %>% 
-    autofit()  
-```
-
 ## Applying Multivariate Modelling Approaches
@@
 #| fig-cap: "Phylogenetic relationships among `r length(tree$tip.label)` *Drosophila* species."
-
+#| include: false
 # Plot the tree
@@
 ```{r, posteriorsz}
+#| include: false
 # We'll now plot the posterior distributions for each species.Obviously our 11 species with known data will have a much tighter distribution.
@@
 TDT curves (and $CT_{max}$) also cannot be used for all organisms due to ethical and conservation considerations.
+Ethical issues could be alleviated by the use of sublethal measures of thermal tolerance like CFL’s (ref), but we do not know how well lethal (TDT curves/$CT_{max}$) and sublethal (CFL’s) measures capture thermal sensitivities/vulnerabilities.
+Identifying whether a correlation exists between lethal and sub-lethal limits will help us understand the mechanisms of vulnerability to heat stress that might lead to population extinction well before mortality is identified.
 Because of these limitations data on thermal sensitivity/heat vulnerability is missing for many species, creating significant knowledge gaps across the tree of life.
@@
 
-While the comparative approach can help fill knowledge gaps it does not entirely solve the issue that TDT curves cannot be estimated for many organisms (logistical and ethical).
-$CT_{max}$ is less logistically challenging but also has ethical limitations.
-Ethical issues could be alleviated by the use of sublethal measures of thermal tolerance like CFL’s (ref), but we do not know how well lethal (TDT curves/$CT_{max}$) and sublethal (CFL’s) measures capture thermal sensitivities/vulnerabilities.
-Identifying whether a correlation exists between lethal and sub-lethal limits will help us understand the mechanisms of vulnerability to heat stress that might lead to population extinction well before mortality is identified.
-And if these traits are correlated different measures of thermal tolerance can be used for different organisms.
-
-#### Application of Multivariate Comparative Approaches using Thermal Tolerance Measures: A case study using *Drosophilids*
+#### Application of Multivariate Phylogenetic Comparative Approaches using Thermal Tolerance Measures: A case study using *Drosophilids*
 
@@
 
-We integrated static lethal limits ($CT_{max}$) collected by @Heerwaarden2021 and XX with dynamic lethal limit measures from TDT curces [i.e., slope ($z^{TDT}$) and intercept ($CT_{max}^{TDT}$)] estimated by XX.
+We integrated static lethal limits ($CT_{max}$) collected by Heerwaarden & Sgrò @Heerwaarden2021 and Kellerman *et al* @Kellerman2012 with dynamic lethal limit measures from TDT curces [i.e., slope ($z^{TDT}$) and intercept ($CT_{max}^{TDT}$)] estimated by Jorgensen *et al* @Jorgensen2019 and .
 We also collected data to estimate TDT curve parameters for eight additonal species following protocols outlined in @Orsted2022.
@@
 
+## Supplementary Methods and Results – Box 1
+
+### Quantifying Thermal Death Time Curves in eight new species of *Drosophila*
+#### Methods
+Thermal death time curves were estimated in eight species of Drosophila: *Drosophila bipectinata*, *Drosophila birchii*, *Drosophila hydei*, *Drosophila kikkawaii*, *Drosophila pseudoananassae*, *Drosophila simulans* and *Drosophila sulfurigaster*.
+All species were collected from the field between 2021 and 2022 by sweep netting over rotten bananas.
+Flies were taken back to the laboratory with isofemale lines established from field inseminated females.
+Mass bred populations were established by combining 20 isofemale lines together, with mass bred population sizes maintained at 20°C 12/12hr light at an approximate population size of 750 individuals.
+Prior to initiation of the experiment flies were placed onto laying pots and 30-40 eggs were collected into each 10-15 replicate vials to control for density.
+On emergence flies were maintained for 4-5 days prior to separating the sexes via CO2 anesthesia.
+Flies were given two days recovery prior to estimation of the thermal death time curves, with TDT curves estimated on on female flies.
+
+For each species thermal death time curves were estimated at five temperatures (35, 36, 37.5, 39 and 40°C), with 11-20 replicates per temperature.
+Flies were placed into a sealed 5mL glass vial and placed into a water set to one of the above temperatures.
+Flies were then monitored continuously and the time at which they knocked down from heat stress and ceased all movement was recorded.
+For two species *D.
+simulans* and *D.
+hydei* we were not able to get all the knockdown time data at 35 °C due to the length of time of the experiment.
+When the experiment was terminated around 50% of *D.
+simulans* had knockdown while not a single individual of *D.
+hydei* had knockdown (although flies were beginning to show signs of heat stress).
+After considering the data with and without this temperature (individuals not knockdown were scored as the time at which the experiment was terminated) we decided to include it.
+By including the data we are likely biasing towards a shallower slope but excluding the data we believe would create a stronger bias in the opposite direction.
+
+#### Results
+
+We analysed the thermal death time curves for 8 new species of *Drosophila* using the same methods as @Jorgensen2019, using linear models with log 10 coma time as the response and assay temperature as the predcitor [@fig-tdt].
+We then used the model to estimate the slope and intercept ($CT_{max}$) for each species [@tbl-1].
+
+```{r, newspp}
+#| label: newspp
+#| echo: false
+#| message: false
+#| include: false
+
+# Load the data
+  spp <- read_csv(here::here("data", "tdt_curves.csv"))
+
+# Calcluate the TDT
+  tdt <- split(spp, spp$group)
+
+# Calculations
+  txt <- plyr::ldply(lapply(tdt, function(x) tdt_calc(data = x)))
+
+```
+
+```{r, fig-tdt}
+#| label: fig-tdt
+#| fig-cap: TDT curves for each species. The x-axis is temperature and the y-axis is the log10 of the time to coma. 
+#| message: false
+#| echo: false
+
+# Lets have a look at the data
+  #ggplot2::ggplot(data, ggplot2::aes(x = assay_temp, y = log10(t_coma), colour = group)) + ggplot2::geom_point() + #ggplot2::facet_wrap(~group, nrow = 2, ncol = 4) +ggplot2::labs(x = "Temperature", y = "log 10 Coma Time") + #ggplot2::theme_bw()
+
+# Fecet wrap doesn't seem to want to work
+  par(mfrow=c(2,4))
+  invisible(lapply(tdt, function(x) plot(log10(x$t_coma) ~ x$assay_temp,  main= unique(x$group), xlab = "Temperature", ylab = "log 10 Coma Time", pch = 16, col = "black")))
+```
+
+```{r, tbl-1}
+#| label: tbl-1
+#| tbl-cap: Estimated parameters for each of the 8 new species. 
+labels <- c("Species", "Slope", "CTMax")
+
+flextable(txt) %>% 
+    align(align = "center", part = "all") %>% 
+    font(fontname="Times", part = "body") %>% 
+    font(fontname="Times", part = "header")  %>%  
+    flextable::compose(i = 1, j = 1:3, part = "header", value= as_paragraph(as_b(labels))) %>% 
+    autofit()  
+```
+
 ## References

### 2023-09-27 ms/ms.qmd
Remove within species We really don't have a lot

@@
 crossref:
-  fig-title: '**Figure**'
+  fig-title: 'Figure'
   fig-labels: arabic
@@
 TDT curves (and $CT_{max}$) also cannot be used for all organisms due to ethical and conservation considerations.
-Ethical issues could be alleviated by the use of sublethal measures of thermal tolerance like CFL’s (ref), but we do not know how well lethal (TDT curves/$CT_{max}$) and sublethal (CFL’s) measures capture thermal sensitivities/vulnerabilities.
+Ethical issues could be alleviated by the use of sublethal measures of thermal tolerance like CFL’s [@Wang2022], but we do not know how well lethal (TDT curves/$CT_{max}$) and sublethal (CFL’s) measures capture thermal sensitivities/vulnerabilities.
 Identifying whether a correlation exists between lethal and sub-lethal limits will help us understand the mechanisms of vulnerability to heat stress that might lead to population extinction well before mortality is identified.
@@
 
-Phylogenetic comparative approaches could be leveraged to help fill these knowledge gaps because closely related species are often more likely to share similar traits owing to their evolutionary history (phylogenetic niche conservatism) (ref).
-If heat tolerance is linked to the phylogeny, shared characteristics that make species more or less heat tolerant will not be randomly distributed across the phylogeny but biased by species groups (ref).
-We can can use evolutionary relatedness to predict species thermal sensitivities and such insights can be used to inform policy and conservation related decisions making for species with little to no data.
+Phylogenetic comparative approaches could be leveraged to help fill these knowledge gaps because closely related species are often more likely to share similar traits owing to their evolutionary history (phylogenetic niche conservatism) [@Losos2008].
+If heat tolerance is linked to the phylogeny, shared characteristics that make species more or less heat tolerant will not be randomly distributed across the phylogeny but biased by species groups [@McKinney1997].
+We can use evolutionary relatedness to predict species thermal sensitivities and such insights can be used to inform policy and conservation related decisions making for species with little to no data.
 
@@
 
-We integrated static lethal limits ($CT_{max}$) collected by Heerwaarden & Sgrò @Heerwaarden2021 and Kellerman *et al* @Kellerman2012 with dynamic lethal limit measures from TDT curces [i.e., slope ($z^{TDT}$) and intercept ($CT_{max}^{TDT}$)] estimated by Jorgensen *et al* @Jorgensen2019 and .
-We also collected data to estimate TDT curve parameters for eight additonal species following protocols outlined in @Orsted2022.
+We integrated static lethal limits ($CT_{max}$) collected by Heerwaarden & Sgrò @Heerwaarden2021 and Kellerman *et al* @Kellerman2012 with dynamic lethal limit measures from TDT curces [i.e., slope ($z^{TDT}$) and intercept ($CT_{max}^{TDT}$)] estimated by Jorgensen *et al* @Jorgensen2019 and Rezende *et al* @Rezende2014.
+We also collected data to estimate TDT curve parameters for eight additonal species following protocols outlined in Jorgensen *et al* @Orsted2022 and Rezende *et al* @Rezende2014 (See *Supplementary Materials*).
 Critical fertility limits for *Drosophilids* were compiled from @Parratt2021 and @Heerwaarden2021.
@@
  Models including phylogeny were better supported than models without phylogeny ($\Delta DIC$ = `r mcmcglmm_model_z_trans_nophylo$DIC - mcmcglmm_model_z_trans$DIC`), with all traits exhibiting a moderate to strong phylogenetic signal indicating evolutionary conservatism in heat tolerance (Phylogenetic $H^2$: $CT_{max}$ – `r posterior.mode(h2_ctmax)`, 95% CI: `r HPDinterval(h2_ctmax)[,"lower"]` to `r HPDinterval(h2_ctmax)[,"upper"]`; $z^{TDT}$ – `r posterior.mode(h2_z)`, 95% CI: `r HPDinterval(h2_z)[,"lower"]` to `r HPDinterval(h2_z)[,"upper"]`; CFLs – `r posterior.mode(h2_cfls)`, 95% CI: `r HPDinterval(h2_cfls)[,"lower"]` to `r HPDinterval(h2_cfls)[,"upper"]`).
-While the model also suggested that there was within species variation, missing data and lack of replicate species populations make conclusions premature ($I^2_{species}$: $CT_{max}$ – `r posterior.mode(spp_ctmax)`, 95% CI: `r HPDinterval(spp_ctmax)[,"lower"]` to `r HPDinterval(spp_ctmax)[,"upper"]`; $z^{TDT}$ – `r posterior.mode(spp_z)`, 95% CI: `r HPDinterval(spp_z)[,"lower"]` to `r HPDinterval(spp_z)[,"upper"]`; CFLs – `r posterior.mode(spp_cfls)`, 95% CI: `r HPDinterval(spp_cfls)[,"lower"]` to `r HPDinterval(spp_cfls)[,"upper"]`).
 
@@
 
-We analysed the thermal death time curves for 8 new species of *Drosophila* using the same methods as @Jorgensen2019, using linear models with log 10 coma time as the response and assay temperature as the predcitor [@fig-tdt].
-We then used the model to estimate the slope and intercept ($CT_{max}$) for each species [@tbl-1].
+We analysed the thermal death time curves for 8 new species of *Drosophila* using the same methods as @Jorgensen2019, using linear models with log 10 coma time as the response and assay temperature as the predcitor (@fig-tdt).
+We then used the model to estimate the slope and intercept ($CT_{max}$) for each species (Table 1).
 

### 2023-09-27 ms/ms.qmd
Updated Box 1.

@@
 
-Phylogenetic comparative approaches could be leveraged to help fill these knowledge gaps because closely related species are often more likely to share similar traits owing to their evolutionary history (phylogenetic niche conservatism) [@Losos2008].
+Phylogenetic comparative approaches could be leveraged to help fill these knowledge gaps because closely related species are often more likely to share similar traits owing to their shared evolutionary history (phylogenetic niche conservatism) [@Losos2008].
 If heat tolerance is linked to the phylogeny, shared characteristics that make species more or less heat tolerant will not be randomly distributed across the phylogeny but biased by species groups [@McKinney1997].
@@
 ##### Evolutionary relationships can help understand thermal tolerance in species lacking data
-The model predicted known data reasonably well for most traits (correlation between known values and values estimated from the model: $z^{TDT}$ – `r cor.test(check_z$trait, check_z$est)$estimate`, 95% CI: `r cor.test(check_z$trait, check_z$est)$conf.int[1]` to `r cor.test(check_z$trait, check_z$est)$conf.int[2]`, p < 0.001; $CT_{max}$ – `r cor.test(check_ctmax$trait, check_ctmax$est)$estimate`, 95% CI: `r cor.test(check_ctmax$trait, check_ctmax$est)$conf.int[1]` to `r cor.test(check_ctmax$trait, check_ctmax$est)$conf.int[2]`, p < 0.001; CFLs – `r cor.test(check_cfls$trait, check_cfls$est)$estimate`, 95% CI: `r cor.test(check_cfls$trait, check_cfls$est)$conf.int[1]` to `r cor.test(check_cfls$trait, check_cfls$est)$conf.int[2]`, p < 0.001), suggesting that even this basic model could be used to provide sensible quantitative measures for *Drosophila* species without heat tolerance estimates.
+The model predicted known data reasonably well for most traits (correlation between known values and values estimated from the model: $z^{TDT}$ – `r cor.test(check_z$trait, check_z$est)$estimate`, 95% CI: `r cor.test(check_z$trait, check_z$est)$conf.int[1]` to `r cor.test(check_z$trait, check_z$est)$conf.int[2]`, p < 0.001, @fig-treesub; $CT_{max}$ – `r cor.test(check_ctmax$trait, check_ctmax$est)$estimate`, 95% CI: `r cor.test(check_ctmax$trait, check_ctmax$est)$conf.int[1]` to `r cor.test(check_ctmax$trait, check_ctmax$est)$conf.int[2]`, p < 0.001; CFLs – `r cor.test(check_cfls$trait, check_cfls$est)$estimate`, 95% CI: `r cor.test(check_cfls$trait, check_cfls$est)$conf.int[1]` to `r cor.test(check_cfls$trait, check_cfls$est)$conf.int[2]`, p < 0.001), suggesting that even this basic model could be used to provide sensible quantitative measures for *Drosophila* species without heat tolerance estimates.
 While modelling approaches such as these are not a replacement for actual empirical data, they could be used to provide informed judgements on thermal tolerance in the absence of such information, but should be tested (using cross-validation) and refined when necessary.
@@
 ##### Static lethal limits are evolutionarily correlated with sub-lethal limits in *Drosophilids*
-Interestingly, there was a reasonably strong phylogenetic correlation between $CT_{max}$ and critical fertility limits in *Drosophilids* (r = `r posterior.mode(phylocor_ctmax_cfls)`, 95% CI: `r HPDinterval(phylocor_ctmax_cfls)[,"lower"]` to `r HPDinterval(phylocor_ctmax_cfls)[,"upper"]`), but not between $CT_{max}$, CFLs and the slope from TDT curves (cor($CT_{max}$, $z^{TDT}$) = `r posterior.mode(phylocor_ctmax_tdt)`, 95% CI: `r HPDinterval(phylocor_ctmax_tdt)[,"lower"]` to `r HPDinterval(phylocor_ctmax_tdt)[,"upper"]`; cor(CFLs, $z^{TDT}$) = `r posterior.mode(phylocor_tdt_cfls)`, 95% CI: `r HPDinterval(phylocor_tdt_cfls)[,"lower"]` to `r HPDinterval(phylocor_tdt_cfls)[,"upper"]`).
+Interestingly, there was a reasonably strong phylogenetic correlation between $CT_{max}$ and critical fertility limits in *Drosophilids* (r = `r posterior.mode(phylocor_ctmax_cfls)`, 95% CI: `r HPDinterval(phylocor_ctmax_cfls)[,"lower"]` to `r HPDinterval(phylocor_ctmax_cfls)[,"upper"]`, @fig-treesub), but not between $CT_{max}$, CFLs and the slope from TDT curves (cor($CT_{max}$, $z^{TDT}$) = `r posterior.mode(phylocor_ctmax_tdt)`, 95% CI: `r HPDinterval(phylocor_ctmax_tdt)[,"lower"]` to `r HPDinterval(phylocor_ctmax_tdt)[,"upper"]`; cor(CFLs, $z^{TDT}$) = `r posterior.mode(phylocor_tdt_cfls)`, 95% CI: `r HPDinterval(phylocor_tdt_cfls)[,"lower"]` to `r HPDinterval(phylocor_tdt_cfls)[,"upper"]`).
 These results suggest that sub-lethal measures, such as CFLs strongly coincide with a species' upper lethal limit.
@@
 
+data_ct_cfts <- data  %>% group_by(species)  %>% summarise(ctmax_m = mean(ctmax),
+                                                          cfts_m = mean(cfts),
+                                                   tdt_z_m = mean(tdt_z))  %>% data.frame()
+
 rerun = FALSE
@@
 
-  layout.matrix <- matrix(c(1, 1, 1, 2,3,4), nrow = 3, ncol = 2) # Phylogeny takes up 3 panels (plot = 1) and the others take up a single panel.
+  layout.matrix <- matrix(c(1, 1,2,3), nrow = 2, ncol = 2) # Phylogeny takes up 3 panels (plot = 1) and the others take up a single panel.
 
@@
   abline(a = 0, b = 1, col = "red")
+
+  # PLot 3
+    plot(ctmax_m ~ cfts_m, data = data_ct_cfts, ylab = "Critcial Thermal Maximum (CTmax)", xlab = "Crictial Fertility Limits (CFLs)", pch = 16, cex = 3, cex.lab = 1.4)
+
   dev.off()
@@
 
-We analysed the thermal death time curves for 8 new species of *Drosophila* using the same methods as @Jorgensen2019, using linear models with log 10 coma time as the response and assay temperature as the predcitor (@fig-tdt).
+We analysed the thermal death time curves for 8 new species of *Drosophila* using the same methods as @Jorgensen2019, using linear models with log 10 coma time as the response and assay temperature as the predictor (@fig-tdt).
 We then used the model to estimate the slope and intercept ($CT_{max}$) for each species (Table 1).

### 2023-09-28 ms/ms.qmd
Add some pdf optoons

@@
     reference-doc: ../bib/template.docx
+  pdf: 
+    documentclass: report
+    margin-left: 30mm
+    margin-right: 30mm
 editor_options: 
